### 2022-01-27 sample-sigplan.tex
Initial Overleaf Import

%%
%% This is file `sample-sigplan.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigplan')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigplan.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigplan,screen]{acmart}
%% NOTE that a single column version is required for 
%% submission and peer review. This can be done by changing
%% the \doucmentclass[...]{acmart} in this template to 
%% \documentclass[manuscript,screen,review]{acmart}
%% 
%% To ensure 100% compatibility, please check the white list of
%% approved LaTeX packages to be used with the Master Article Template at
%% https://www.acm.org/publications/taps/whitelist-of-latex-packages 
%% before creating your document. The white list page provides 
%% information on how to submit additional LaTeX packages for 
%% review and adoption.
%% Fonts used in the template cannot be substituted; margin 
%% adjustments are not allowed.
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Conference acronym 'XX]{Make sure to enter the correct
  conference title from your rights confirmation emai}{June 03--05,
  2018}{Woodstock, NY}
%
%  Uncomment \acmBooktitle if th title of the proceedings is different
%  from ``Proceedings of ...''!
%
%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%  June 03--05, 2018, Woodstock, NY} 
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{The Name of the Title is Hope}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Ben Trovato}
\authornote{Both authors contributed equally to this research.}
\email{trovato@corporation.com}
\orcid{1234-5678-9012}
\author{G.K.M. Tobin}
\authornotemark[1]
\email{webmaster@marysville-ohio.com}
\affiliation{%
  \institution{Institute for Clarity in Documentation}
  \streetaddress{P.O. Box 1212}
  \city{Dublin}
  \state{Ohio}
  \country{USA}
  \postcode{43017-6221}
}

\author{Lars Th{\o}rv{\"a}ld}
\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
  \city{Hekla}
  \country{Iceland}}
\email{larst@affiliation.org}

\author{Valerie B\'eranger}
\affiliation{%
  \institution{Inria Paris-Rocquencourt}
  \city{Rocquencourt}
  \country{France}
}

\author{Aparna Patel}
\affiliation{%
 \institution{Rajiv Gandhi University}
 \streetaddress{Rono-Hills}
 \city{Doimukh}
 \state{Arunachal Pradesh}
 \country{India}}

\author{Huifen Chan}
\affiliation{%
  \institution{Tsinghua University}
  \streetaddress{30 Shuangqing Rd}
  \city{Haidian Qu}
  \state{Beijing Shi}
  \country{China}}

\author{Charles Palmer}
\affiliation{%
  \institution{Palmer Research Laboratories}
  \streetaddress{8600 Datapoint Drive}
  \city{San Antonio}
  \state{Texas}
  \country{USA}
  \postcode{78229}}
\email{cpalmer@prl.com}

\author{John Smith}
\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
  \city{Hekla}
  \country{Iceland}}
\email{jsmith@affiliation.org}

\author{Julius P. Kumquat}
\affiliation{%
  \institution{The Kumquat Consortium}
  \city{New York}
  \country{USA}}
\email{jpkumquat@consortium.net}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Trovato and Tobin, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  A clear and well-documented \LaTeX\ document is presented as an
  article formatted for publication by ACM in a conference proceedings
  or journal publication. Based on the ``acmart'' document class, this
  article presents and explains many of the common variations, as well
  as many of the formatting elements an author may use in the
  preparation of the documentation of their work.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{datasets, neural networks, gaze detection, text tagging}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \includegraphics[width=\textwidth]{sampleteaser}
  \caption{Seattle Mariners at Spring Training, 2010.}
  \Description{Enjoying the baseball game from the third-base
  seats. Ichiro Suzuki preparing to bat.}
  \label{fig:teaser}
\end{teaserfigure}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
ACM's consolidated article template, introduced in 2017, provides a
consistent \LaTeX\ style for use across ACM publications, and
incorporates accessibility and metadata-extraction functionality
necessary for future Digital Library endeavors. Numerous ACM and
SIG-specific \LaTeX\ templates have been examined, and their unique
features incorporated into this single new template.

If you are new to publishing with ACM, this document is a valuable
guide to the process of preparing your work for publication. If you
have published with ACM before, this document provides insight and
instruction into more recent changes to the article template.

The ``\verb|acmart|'' document class can be used to prepare articles
for any ACM publication --- conference or journal, and for any stage
of publication, from review to final ``camera-ready'' copy, to the
author's own version, with {\itshape very} few changes to the source.

\section{Template Overview}
As noted in the introduction, the ``\verb|acmart|'' document class can
be used to prepare many different kinds of documentation --- a
double-blind initial submission of a full-length technical paper, a
two-page SIGGRAPH Emerging Technologies abstract, a ``camera-ready''
journal article, a SIGCHI Extended Abstract, and more --- all by
selecting the appropriate {\itshape template style} and {\itshape
  template parameters}.

This document will explain the major features of the document
class. For further information, the {\itshape \LaTeX\ User's Guide} is
available from
\url{https://www.acm.org/publications/proceedings-template}.

\subsection{Template Styles}

The primary parameter given to the ``\verb|acmart|'' document class is
the {\itshape template style} which corresponds to the kind of publication
or SIG publishing the work. This parameter is enclosed in square
brackets and is a part of the {\verb|documentclass|} command:
\begin{verbatim}
  \documentclass[STYLE]{acmart}
\end{verbatim}

Journals use one of three template styles. All but three ACM journals
use the {\verb|acmsmall|} template style:
\begin{itemize}
\item {\verb|acmsmall|}: The default journal template style.
\item {\verb|acmlarge|}: Used by JOCCH and TAP.
\item {\verb|acmtog|}: Used by TOG.
\end{itemize}

The majority of conference proceedings documentation will use the {\verb|acmconf|} template style.
\begin{itemize}
\item {\verb|acmconf|}: The default proceedings template style.
\item{\verb|sigchi|}: Used for SIGCHI conference articles.
\item{\verb|sigchi-a|}: Used for SIGCHI ``Extended Abstract'' articles.
\item{\verb|sigplan|}: Used for SIGPLAN conference articles.
\end{itemize}

\subsection{Template Parameters}

In addition to specifying the {\itshape template style} to be used in
formatting your work, there are a number of {\itshape template parameters}
which modify some part of the applied template style. A complete list
of these parameters can be found in the {\itshape \LaTeX\ User's Guide.}

Frequently-used parameters, or combinations of parameters, include:
\begin{itemize}
\item {\verb|anonymous,review|}: Suitable for a ``double-blind''
  conference submission. Anonymizes the work and includes line
  numbers. Use with the \verb|\acmSubmissionID| command to print the
  submission's unique ID on each page of the work.
\item{\verb|authorversion|}: Produces a version of the work suitable
  for posting by the author.
\item{\verb|screen|}: Produces colored hyperlinks.
\end{itemize}

This document uses the following string as the first command in the
source file:
\begin{verbatim}
\documentclass[sigplan,screen]{acmart}
\end{verbatim}

\section{Modifications}

Modifying the template --- including but not limited to: adjusting
margins, typeface sizes, line spacing, paragraph and list definitions,
and the use of the \verb|\vspace| command to manually adjust the
vertical spacing between elements of your work --- is not allowed.

{\bfseries Your document will be returned to you for revision if
  modifications are discovered.}

\section{Typefaces}

The ``\verb|acmart|'' document class requires the use of the
``Libertine'' typeface family. Your \TeX\ installation should include
this set of packages. Please do not substitute other typefaces. The
``\verb|lmodern|'' and ``\verb|ltimes|'' packages should not be used,
as they will override the built-in typeface families.

\section{Title Information}

The title of your work should use capital letters appropriately -
\url{https://capitalizemytitle.com/} has useful rules for
capitalization. Use the {\verb|title|} command to define the title of
your work. If your work has a subtitle, define it with the
{\verb|subtitle|} command.  Do not insert line breaks in your title.

If your title is lengthy, you must define a short version to be used
in the page headers, to prevent overlapping text. The \verb|title|
command has a ``short title'' parameter:
\begin{verbatim}
  \title[short title]{full title}
\end{verbatim}

\section{Authors and Affiliations}

Each author must be defined separately for accurate metadata
identification. Multiple authors may share one affiliation. Authors'
names should not be abbreviated; use full first names wherever
possible. Include authors' e-mail addresses whenever possible.

Grouping authors' names or e-mail addresses, or providing an ``e-mail
alias,'' as shown below, is not acceptable:
\begin{verbatim}
  \author{Brooke Aster, David Mehldau}
  \email{dave,judy,steve@university.edu}
  \email{firstname.lastname@phillips.org}
\end{verbatim}

The \verb|authornote| and \verb|authornotemark| commands allow a note
to apply to multiple authors --- for example, if the first two authors
of an article contributed equally to the work.

If your author list is lengthy, you must define a shortened version of
the list of authors to be used in the page headers, to prevent
overlapping text. The following command should be placed just after
the last \verb|\author{}| definition:
\begin{verbatim}
  \renewcommand{\shortauthors}{McCartney, et al.}
\end{verbatim}
Omitting this command will force the use of a concatenated list of all
of the authors' names, which may result in overlapping text in the
page headers.

The article template's documentation, available at
\url{https://www.acm.org/publications/proceedings-template}, has a
complete explanation of these commands and tips for their effective
use.

Note that authors' addresses are mandatory for journal articles.

\section{Rights Information}

Authors of any work published by ACM will need to complete a rights
form. Depending on the kind of work, and the rights management choice
made by the author, this may be copyright transfer, permission,
license, or an OA (open access) agreement.

Regardless of the rights management choice, the author will receive a
copy of the completed rights form once it has been submitted. This
form contains \LaTeX\ commands that must be copied into the source
document. When the document source is compiled, these commands and
their parameters add formatted text to several areas of the final
document:
\begin{itemize}
\item the ``ACM Reference Format'' text on the first page.
\item the ``rights management'' text on the first page.
\item the conference information in the page header(s).
\end{itemize}

Rights information is unique to the work; if you are preparing several
works for an event, make sure to use the correct set of commands with
each of the works.

The ACM Reference Format text is required for all articles over one
page in length, and is optional for one-page articles (abstracts).

\section{CCS Concepts and User-Defined Keywords}

Two elements of the ``acmart'' document class provide powerful
taxonomic tools for you to help readers find your work in an online
search.

The ACM Computing Classification System ---
\url{https://www.acm.org/publications/class-2012} --- is a set of
classifiers and concepts that describe the computing
discipline. Authors can select entries from this classification
system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the
commands to be included in the \LaTeX\ source.

User-defined keywords are a comma-separated list of words and phrases
of the authors' choosing, providing a more flexible way of describing
the research being presented.

CCS concepts and user-defined keywords are required for for all
articles over two pages in length, and are optional for one- and
two-page articles (or abstracts).

\section{Sectioning Commands}

Your work should use standard \LaTeX\ sectioning commands:
\verb|section|, \verb|subsection|, \verb|subsubsection|, and
\verb|paragraph|. They should be numbered; do not remove the numbering
from the commands.

Simulating a sectioning command by setting the first word or words of
a paragraph in boldface or italicized text is {\bfseries not allowed.}

\section{Tables}

The ``\verb|acmart|'' document class includes the ``\verb|booktabs|''
package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
high-quality tables.

Table captions are placed {\itshape above} the table.

Because tables cannot be split across pages, the best placement for
them is typically the top of the page nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and the
table caption.  The contents of the table itself must go in the
\textbf{tabular} environment, to be aligned properly in rows and
columns, with the desired horizontal and vertical rules.  Again,
detailed instructions on \textbf{tabular} material are found in the
\textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table~\ref{tab:freq} is included in the input file; compare the
placement of the table here with the table in the printed output of
this document.

\begin{table}
  \caption{Frequency of Special Characters}
  \label{tab:freq}
  \begin{tabular}{ccl}
    \toprule
    Non-English or Math&Frequency&Comments\\
    \midrule
    \O & 1 in 1,000& For Swedish names\\
    $\pi$ & 1 in 5& Common in math\\
    \$ & 4 in 5 & Used in business\\
    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
  \bottomrule
\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of the page's
live area, use the environment \textbf{table*} to enclose the table's
contents and the table caption.  As with a single-column table, this
wide table will ``float'' to a location deemed more
desirable. Immediately following this sentence is the point at which
Table~\ref{tab:commands} is included in the input file; again, it is
instructive to compare the placement of the table here with the table
in the printed output of this document.

\begin{table*}
  \caption{Some Typical Commands}
  \label{tab:commands}
  \begin{tabular}{ccl}
    \toprule
    Command &A Number & Comments\\
    \midrule
    \texttt{{\char'134}author} & 100& Author \\
    \texttt{{\char'134}table}& 300 & For tables\\
    \texttt{{\char'134}table*}& 400& For wider tables\\
    \bottomrule
  \end{tabular}
\end{table*}

Always use midrule to separate table header rows from data rows, and
use it only for this purpose. This enables assistive technologies to
recognise table headers and support their users in navigating tables
more easily.

\section{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of the three are
discussed in the next sections.

\subsection{Inline (In-text) Equations}
A formula that appears in the running text is called an inline or
in-text formula.  It is produced by the \textbf{math} environment,
which can be invoked with the usual
\texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
and structures, from $\alpha$ to $\omega$, available in
\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
examples of in-text equations in context. Notice how this equation:
\begin{math}
  \lim_{n\rightarrow \infty}x=0
\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsection{Display Equations}
A numbered display equation---one set off by vertical space from the
text and centered horizontally---is produced by the \textbf{equation}
environment. An unnumbered display equation is produced by the
\textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols and
structures available in \LaTeX\@; this section will just give a couple
of examples of display equations in context.  First, consider the
equation, shown as an inline equation above:
\begin{equation}
  \lim_{n\rightarrow \infty}x=0
\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}
  \sum_{i=0}^{\infty} x + 1
\end{displaymath}
and follow it with another numbered equation:
\begin{equation}
  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\section{Figures}

The ``\verb|figure|'' environment should be used for figures. One or
more images can be placed within a figure. If your figure contains
third-party material, you must clearly identify it as such, as shown
in the example below.
\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{sample-franklin}
  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
    Ewing, Inc. [Public domain], via Wikimedia
    Commons. (\url{https://goo.gl/VLCRBB}).}
  \Description{A woman and a girl in white dresses sit in an open car.}
\end{figure}

Your figures should contain a caption which describes the figure to
the reader.

Figure captions are placed {\itshape below} the figure.

Every figure should also have a figure description unless it is purely
decorative. These descriptions convey what’s in the image to someone
who cannot see it. They are also used by search engine crawlers for
indexing images, and when images cannot be loaded.

A figure description must be unformatted plain text less than 2000
characters long (including spaces).  {\bfseries Figure descriptions
  should not repeat the figure caption – their purpose is to capture
  important information that is not already provided in the caption or
  the main text of the paper.} For figures that convey important and
complex new information, a short text description may not be
adequate. More complex alternative descriptions can be placed in an
appendix and referenced in a short figure description. For example,
provide a data table capturing the information in a bar chart, or a
structured list representing a graph.  For additional information
regarding how best to write figure descriptions and why doing this is
so important, please see
\url{https://www.acm.org/publications/taps/describing-figures/}.

\subsection{The ``Teaser Figure''}

A ``teaser figure'' is an image, or set of images in one figure, that
are placed after all author and affiliation information, and before
the body of the article, spanning the page. If you wish to have such a
figure in your article, place the command immediately before the
\verb|\maketitle| command:
\begin{verbatim}
  \begin{teaserfigure}
    \includegraphics[width=\textwidth]{sampleteaser}
    \caption{figure caption}
    \Description{figure description}
  \end{teaserfigure}
\end{verbatim}

\section{Citations and Bibliographies}

The use of \BibTeX\ for the preparation and formatting of one's
references is strongly recommended. Authors' names should be complete
--- use full first names (``Donald E. Knuth'') not initials
(``D. E. Knuth'') --- and the salient identifying features of a
reference should be included: title, year, volume, number, pages,
article DOI, etc.

The bibliography is included in your source document with these two
commands, placed just before the \verb|\end{document}| command:
\begin{verbatim}
  \bibliographystyle{ACM-Reference-Format}
  \bibliography{bibfile}
\end{verbatim}
where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
suffix, of the \BibTeX\ file.

Citations and references are numbered by default. A small number of
ACM publications have citations and references formatted in the
``author year'' style; for these exceptions, please include this
command in the {\bfseries preamble} (before the command
``\verb|\begin{document}|'') of your \LaTeX\ source:
\begin{verbatim}
  \citestyle{acmauthoryear}
\end{verbatim}

  Some examples.  A paginated journal article \cite{Abril07}, an
  enumerated journal article \cite{Cohen07}, a reference to an entire
  issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
  monograph/whole book in a series (see 2a in spec. document)
  \cite{Harel79}, a divisible-book such as an anthology or compilation
  \cite{Editor00} followed by the same example, however we only output
  the series if the volume number is given \cite{Editor00a} (so
  Editor00a's series should NOT be present since it has no vol. no.),
  a chapter in a divisible book \cite{Spector90}, a chapter in a
  divisible book in a series \cite{Douglass98}, a multi-volume work as
  book \cite{Knuth97}, a couple of articles in a proceedings (of a
  conference, symposium, workshop for example) (paginated proceedings
  article) \cite{Andler79, Hagerup1993}, a proceedings article with
  all possible elements \cite{Smith10}, an example of an enumerated
  proceedings article \cite{VanGundy07}, an informally published work
  \cite{Harel78}, a couple of preprints \cite{Bornmann2019,
    AnzarootPBM14}, a doctoral dissertation \cite{Clarkson85}, a
  master's thesis: \cite{anisi03}, an online document / world wide web
  resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game
  (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05}
  and (Case 3) a patent \cite{JoeScientist001}, work accepted for
  publication \cite{rous08}, 'YYYYb'-test for prolific author
  \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
  contain 'duplicate' DOI and URLs (some SIAM articles)
  \cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
  multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
  couple of citations with DOIs:
  \cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
  citations: \cite{TUGInstmem, Thornburg01, CTANacmart}. Artifacts:
  \cite{R} and \cite{UMassCitations}.

\section{Acknowledgments}

Identification of funding sources and other support, and thanks to
individuals and groups that assisted in the research and the
preparation of the work should be included in an acknowledgment
section, which is placed just before the reference section in your
document.

This section has a special environment:
\begin{verbatim}
  \begin{acks}
  ...
  \end{acks}
\end{verbatim}
so that the information contained therein can be more easily collected
during the article metadata extraction phase, and to ensure
consistency in the spelling of the section heading.

Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.

\section{Appendices}

If your work needs an appendix, add it before the
``\verb|\end{document}|'' command at the conclusion of your source
document.

Start the appendix with the ``\verb|appendix|'' command:
\begin{verbatim}
  \appendix
\end{verbatim}
and note that in the appendix, sections are lettered, not
numbered. This document has two appendices, demonstrating the section
and subsection identification method.

\section{SIGCHI Extended Abstracts}

The ``\verb|sigchi-a|'' template style (available only in \LaTeX\ and
not in Word) produces a landscape-orientation formatted article, with
a wide left margin. Three environments are available for use with the
``\verb|sigchi-a|'' template style, and produce formatted output in
the margin:
\begin{itemize}
\item {\verb|sidebar|}:  Place formatted text in the margin.
\item {\verb|marginfigure|}: Place a figure in the margin.
\item {\verb|margintable|}: Place a table in the margin.
\end{itemize}

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
To Robert, for the bagels and explaining CMYK and color spaces.
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\section{Research Methods}

\subsection{Part One}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
lacinia dolor. Integer ultricies commodo sem nec semper.

\subsection{Part Two}

Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
eros. Vivamus non purus placerat, scelerisque diam eu, cursus
ante. Etiam aliquam tortor auctor efficitur mattis.

\section{Online Resources}

Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
enim maximus. Vestibulum gravida massa ut felis suscipit
congue. Quisque mattis elit a risus ultrices commodo venenatis eget
dui. Etiam sagittis eleifend elementum.

Nam interdum magna at lectus dignissim, ac dignissim lorem
rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
massa et mattis lacinia.

\end{document}
\endinput
%%
%% End of file `sample-sigplan.tex'.

### 2022-01-28 paper-capstone-earthquake.tex
cleanup paper, added refernces, benchmark section, todo command, todos

\documentclass[sigplan,screen]{acmart}
\usepackage{todonotes}

\newcommand{\TODO}[1]{\todo[inline]{#1}}

\setcopyright{acmcopyright}
\copyrightyear{2022}
\acmYear{2022}
\acmDOI{XXXXXXX.XXXXXXX}

\acmConference[Conference acronym 'XX]{Make sure to enter the correct
  conference title from your rights confirmation emai}{June 03--05,
  2018}{Woodstock, NY}
%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%  June 03--05, 2018, Woodstock, NY} 
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}
%%\acmSubmissionID{123-A56-BU3}
%%\citestyle{acmauthoryear}

\begin{document}

\title{MLCommons Earthquake Science Benchmark}

\author{Thomas Butler}
\email{trovato@corporation.com}
\author{G.K.M. Tobin}
\email{webmaster@marysville-ohio.com}
\affiliation{%
  \institution{University of Virginia}
  \streetaddress{TBD}
  \city{Charlotsville, VA}
  \country{USA}}

\author{Robert Knuuti}
\affiliation{%
  \institution{University of Virginia}
  \streetaddress{TBD}
  \city{Charoletteville}
  \country{USA}}
\email{uqq5zz@virginia.edu}

\author{Jake Kolessar}
\affiliation{%
  \institution{University of Virginia}
  \streetaddress{TBD}
  \city{Charoletteville}
  \country{United States}}
\email{jakekolessar@gmail.com}

\author{Geoffrey C. Fox}
\affiliation{%
  \institution{University of Virginia}
  \streetaddress{TBD}
  \city{Charlotsville, VA}
  \country{USA}}
\email{gcfexchange@gmail.com}

\author{Gregor von Laszewski}
\orcid{0000-0001-9558-179X}
\affiliation{%
  \institution{University of Virginia}
  \streetaddress{TBD}
  \city{Charlotsville, VA}
  \country{USA}}
\email{laszewski@gmail.com}

\renewcommand{\shortauthors}{Buttler, Knuuti, Kolesar, Fox, von Laszewski}

\begin{abstract}
TBD
\end{abstract}

%% see http://dl.acm.org/ccs.cfm.
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{MLCommons, Science AI Benchmark, Earthquake, Deep Learning}

\maketitle

\section{Introduction}

\subsection{MLCommons}

\subsection{Earthquake prediction}

Things to do: 
\begin{itemize}
    %\item add presentation as cite,
    \item Review IEEE Presentation \cite{fox2022aiforscience}
    \item Review MLCommons \cite{www-mlcommons} 
    \item Use jabref \cite{www-jabrefg-org} for citation management
    \item Review Fox paper \cite{fox2021earthquake}
    \item Frequently check github \cite{www-mlcommons-eathquake}
    \item Read up on TFT \cite{www-onnen2021}
    \item Become familiar with the Attention Paper \cite{vaswani2017attention}
    \item Learn how to selfdeploy jupyterlab \cite{www-jupyterlab}
    \TODO{Is this the same as jupytext?}

    \item Begin to learn about papermill \cite{www-papermill} \TODO{Rivanna's core configurations might be a bit limited on this front.  They do provide multiple versions of software (such as py2.7 and py3.8), but we may need to get permission to do something in userspace if we need to be very specific on a version of python.  Do we want to look into using tox, conda, or leave this up to the container ecosystem to solve?}
    \TODO{Gregor: add how to use conda and modules to switch python versions}
    \item virtualenv on rivanna for a particular python version.
    \item depends on Tensorflow
    \TODO{Below this line are objectives or targets to take the current modeling solution and mature it for other platforms / ecosystems.  Rivanna uses Lua's lmod ecosystem for jailing a process, and anaconda uses solved environments for dependencies that extend beyond just python modules.}
    \item Familiarize yourself with Rivanna's modules \cite{www-modules}  and conda \cite{www-conda} environments.
    \item pytorch \cite{www-pytorch}
    \TODO{There is interest in comparing pytorch and tensoflow}
    \item horovod \cite{www-horovod}
    \TODO{MLCommons project that can target a few platforms using a YAML contract.  Once we solve the target environment, we can likely target porting this way.}
    \TODO{Also, good news is Rivanna has a Singularity module  as an lmod (3.7.1), so we might be able to use that as a targeted runner without having to setup too much additional scaffolding.
    And if things work right, we could use the k8s module to do local deployments if we wanted to test out alternate execution paths. }
    \item mlcube \cite{www-mlcube} 
\end{itemize}

%\begin{figure}[h]
%  \centering
%  \includegraphics[width=\linewidth]{sample-franklin}
%  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
%    Ewing, Inc. [Public domain], via Wikimedia
%    Commons. (\url{https://goo.gl/VLCRBB}).}
%  \Description{A woman and a girl in white dresses sit in an open car.}
% \end{figure}

\section{Benchmarks}

\subsection{Hardware}

\subsection{University of Virginia Rivanna}
\label{id:uva}

UVA, Rivanna, A100, RTX3090, others may not have enough memory. 
\TODO{description of Rivanna on web page is not accurate.}

\subsection{Rutherfordlab ???}
\label{id:ral}

Rutherfordlab (RAL)

\subsection{Summit}
\label{id:ornl}

Oak Ridge National Labratory (ORNL), Summit

\subsection{NERSC ???}
\label{id:nersc}

National Energy Research Scientific Computing Center (NERSC), ...

\subsection{Aurora}
\label{id:anl}

Argonne National Laboratory (ANL), ...

\subsection{SDSC}
\label{id:sdcs}

San Diego Super Computing Center, ...

\subsection {Nvidia DGX Workstation}

\label{id:dgx} DGX Fox \cite{www-dgx-station-a100}

\begin{table*}[htb]
    \caption{Overview of compute resources.}
    \label{tab:my_label}
    \centering
    \begin{tabular}{|r|l|l|r|l|l|l|}
        \hline
        Section & Organization & Machine  & Processors & GPUs & GPU Type  & Commissioned \\ 
        \hline
        \hline
         \ref{id:uva}   & UVA       & Rivanna   &            &      &           & \\
         \ref{id:ral}   & RAL       &           &            &      &           & \\
         \ref{id:ornl}  & ORNL      &  Summit   &            &      &           & \\
         \ref{id:nersc} & NERSC     &           &            &      &           & \\
         \ref{id:anl}   & ANL       &           &            &      &           & \\
         \ref{id:sdsc}  & SDSC      &           &            &      &           & \\
         \ref{id:dgx}   & DGX Staion A100   &           & 1          &  AMD EPYC 7742 64-Core Processor   & A100 80GB  & Fox will know\\
         \hline
    \end{tabular}
\end{table*}

\begin{acks}
\TODO{Geoffrey: please add acknowledgement for funding, other acks}
\end{acks}

\bibliographystyle{ACM-Reference-Format}
\bibliography{references,sample-base}

\clearpage

\appendix

\section{Manuals Developed}

The following manuals have been developed by the team:

\begin{itemize}
    \item Introduction to Python \cite{las-intro-python}
\end{itemize}

\section{Todo}

\listoftodos

Robert: I am moving the rough notes into our github repo and made a

\begin{itemize}
\item \href{https://github.com/cybertraining-dsc/capstone-eartquake/blob/main/TODO.md}{TODO.md} 
\end{itemize}

file that we can use to build a simple list of things we want to do.  I'll leave them in the overleaf for now, but I feel like that's not a good home for that type of information given that it's mean to be the paper to submit on the subject.

\TODO{Not sure if we want to talk about building tasks / kanban boards to track the different parts of work we'll be doing each week.  Additionally, it might be good to decide on some core roles to make sure things don't fall through the cracks (roles like recorder, scheduler, reviewer, et al...); the non-technical stuff for now, and then we can get into the more specific things once we pass the initial knowledge rampup}

\end{document}
\endinput
%%
%% End of file `sample-sigplan.tex'.

### 2022-01-28 sample.tex
cleanup paper, added refernces, benchmark section, todo command, todos

--- 2022-01-27 sample-sigplan.tex
@@
 \setcopyright{acmcopyright}
-\copyrightyear{2018}
-\acmYear{2018}
+\copyrightyear{2022}
+\acmYear{2022}
 \acmDOI{XXXXXXX.XXXXXXX}
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{The Name of the Title is Hope}
+\title{MLCommons Earthquake Science Benchmark}
 
@@
 %% used to denote shared contribution to the research.
-\author{Ben Trovato}
+\author{Thomas Butler}
 \authornote{Both authors contributed equally to this research.}
@@
 
-\author{Lars Th{\o}rv{\"a}ld}
+\author{Robert Knuuti}
 \affiliation{%
-  \institution{The Th{\o}rv{\"a}ld Group}
-  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
-  \city{Hekla}
-  \country{Iceland}}
-\email{larst@affiliation.org}
+  \institution{University of Virginia}
+  \streetaddress{???}
+  \city{Charoletteville}
+  \country{United States}}
+\email{uqq5zz@virginia.edu}
 
@@
 
-\author{Aparna Patel}
+\author{Geoffrey C. Fox}
 \affiliation{%
- \institution{Rajiv Gandhi University}
- \streetaddress{Rono-Hills}
- \city{Doimukh}
- \state{Arunachal Pradesh}
- \country{India}}
+  \institution{University of Virginia}
+  \streetaddress{TBD}
+  \city{Charlotsville, VA}
+  \country{USA}}
+\email{gcfexchange@gmail.com}
 
-\author{Huifen Chan}
+\author{Gregor von Laszewski}
 \affiliation{%
-  \institution{Tsinghua University}
-  \streetaddress{30 Shuangqing Rd}
-  \city{Haidian Qu}
-  \state{Beijing Shi}
-  \country{China}}
-
-\author{Charles Palmer}
-\affiliation{%
-  \institution{Palmer Research Laboratories}
-  \streetaddress{8600 Datapoint Drive}
-  \city{San Antonio}
-  \state{Texas}
-  \country{USA}
-  \postcode{78229}}
-\email{cpalmer@prl.com}
-
-\author{John Smith}
-\affiliation{%
-  \institution{The Th{\o}rv{\"a}ld Group}
-  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
-  \city{Hekla}
-  \country{Iceland}}
-\email{jsmith@affiliation.org}
-
-\author{Julius P. Kumquat}
-\affiliation{%
-  \institution{The Kumquat Consortium}
-  \city{New York}
+  \institution{University of Virginia}
+  \streetaddress{TBD}
+  \city{Charlotsville, VA}
   \country{USA}}
-\email{jpkumquat@consortium.net}
+\email{laszewski@gmail.com}
 
@@
 %% page.
-\begin{teaserfigure}
-  \includegraphics[width=\textwidth]{sampleteaser}
-  \caption{Seattle Mariners at Spring Training, 2010.}
-  \Description{Enjoying the baseball game from the third-base
-  seats. Ichiro Suzuki preparing to bat.}
-  \label{fig:teaser}
-\end{teaserfigure}
+%\begin{teaserfigure}
+%  \includegraphics[width=\textwidth]{sampleteaser}
+%  \caption{Seattle Mariners at Spring Training, 2010.}
+%  \Description{Enjoying the baseball game from the third-base
+%  seats. Ichiro Suzuki preparing to bat.}
+%  \label{fig:teaser}
+%\end{teaserfigure}
 
@@
 \section{Introduction}
+
+\subsection{MLCommons}
+
+\subsection{Earthquake prediction}
+
+Things to do: 
+\begin{itemize}
+    %\item add presentation as cite,
+    \item Review IEEE Presentation \cite{fox2022aiforscience}
+    \item Review MLCommons \cite{mlcommonswebsite} 
+    \item Use jabref \cite{www-jabrefg-org} for citation management
+    \item Review Fox paper \cite{fox2021earthquake}
+    \item Frequently check github \cite{foxdengit2022}
+    \item Read up on TFT \cite{onnen2021}
+    \item Become familiar with the Attention Paper \cite{vaswani2017attention}
+    \item Learn how to selfdeploy jupyterlab \cite{jupyterlab}
+    %% Is this the same as jupytext?
+    \item Begin to learn about papermill (https://github.com/nteract/papermill)
+    %% Rivanna's core configurations might be a bit limited on this front.
+They do provide multiple versions of software (such as py2.7 and py3.8), but we may need to get permission to do something in userspace if we need to be very specific on a version of python.
+Do we want to look into using tox, conda, or leave this up to the container ecosystem to solve?
+    \item virtualenv on rivanna for a particular python version.
+    \item depends on Tensorflow
+    %% Below this line are objectives or targets to take the current modeling solution and mature it for other platforms / ecosystems.
+Rivanna uses Lua's lmod ecosystem for jailing a process, and anaconda uses solved environments for dependencies that extend beyond just python modules.
+    \item Familiarize yourself with Rivanna's modules (https://lmod.readthedocs.org/) and conda (https://docs.conda.io/en/latest/miniconda.html) environments.
+    %% There's interest in pytorch
+    \item pytorch
+    %% This the right link? https://github.com/horovod/horovod
+    \item horovod
+    %% MLCommons project that can target a few platforms using a YAML contract.  Once we solve the target environment, we can likely target porting this way.
+    %% Also, good news is Rivanna has a Singularity module (https://mlcommons.github.io/mlcube/getting-started/) as an lmod (3.7.1), so we might be able to use that as a targeted runner without having to setup too much additional scaffolding.
+    %% And if things work right, we could use the k8s module to do local deployments if we wanted to test out alternate execution paths.
+    \item mlcube
+\end{itemize}
+
+TODO: delete the rest of paper while keeping one image and 1 cite as example.
+
+%% Pulled Rivanna details into github markdown file.
+
+\section{original intro}
+
 ACM's consolidated article template, introduced in 2017, provides a
@@
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{sample-base}
+\bibliography{references,sample-base}
 

### 2022-01-30 paper-capstone-earthquake.tex
recover refernces, remove acm template

@@
-\documentclass[sigplan,screen]{acmart}
+\documentclass[sigplan,screen]{format/acmart}
 \usepackage{todonotes}
+\usepackage{graphicx}
+\usepackage{wrapfig}
 
@@
 
-\setcopyright{acmcopyright}
+%\setcopyright{acmcopyright}
+\setcopyright{Apachee 2.0}
 \copyrightyear{2022}
@@
 
-\acmConference[Conference acronym 'XX]{Make sure to enter the correct
-  conference title from your rights confirmation emai}{June 03--05,
-  2018}{Woodstock, NY}
-%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%  June 03--05, 2018, Woodstock, NY} 
-\acmPrice{15.00}
-\acmISBN{978-1-4503-XXXX-X/18/06}
+\acmConference[University of Virginia Data Science Capstone 2022]{MLCommons Earthquake Science Benchmark}{Jan. 22 -- Aug. ??,
+  2022}{Charlotsville, VA}
+\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
+  June 03--05, 2018, Woodstock, NY} 
+\acmPrice{0.00}
+\acmISBN{TBD}
 %%\acmSubmissionID{123-A56-BU3}
@@
 \author{Robert Knuuti}
+\orcid{0000-0003-2614-0899}
 \affiliation{%
@@
   \country{USA}}
-\email{uqq5zz@virginia.edu}
+\email{robert.knuuti@gmail.com}
 
@@
 \email{laszewski@gmail.com}
+
+\author{Judy Fox}
+\affiliation{%
+  \institution{University of Virginia}
+  \streetaddress{TBD}
+  \city{Charlotsville, VA}
+  \country{USA}}
+\email{cwk9mp@virginia.edu}
 
@@
 
+Several research efforts have been executed over the years to establish a set of leading factors that could be used to predict where an earthquake could occur.
+Not only is this an algorithmically complex problem, but it is also computationally difficult to perform\cite{fox2022aiforscience}.
+The intersection of these two elements create an ideal environment for building benchmarks.
+By establishing a rigorous process to build a repeatable models, we can establish a benchmark to not only measure the performance of high performance supercomputers, but also take steps to building a stronger understanding of what models perform best to predict when and earthquake could occur.
+
+For this study, we review the following models across various hardware platforms to establish this benchmark:
+\begin{itemize}
+    \item pure (vanilla) recurrent neuronetwork
+    \item spacio-temporal science transformer
+    \item temporal fusion transformer
+\end{itemize}
+
 \subsection{MLCommons}
+
+MLCommons is an open engineering organization that seeks to unify the engineering and machine learning communities through system benchmarks, open data, and best practices\ref{www-mlcommons}.
+By using open data, a reference implementation for benchmarking, and executing best practices, you can build the necessary rigor to construct repeatable research to establish a basis when performing future analysis.
+
+This organization will establish the dimensions for what is the minimally viable solution and to establish our comparison criteria across our models and the executions on the different hardware sets.
+
+\TODO{Need to confirm what the criteria of the benchmark submission is.}
 
@@
 Things to do:
+
 \begin{itemize}
-    %\item add presentation as cite,
+    %% \item add presentation as cite,
     \item Review IEEE Presentation \cite{fox2022aiforscience}
@@
 UVA, Rivanna, A100, RTX3090, others may not have enough memory. 
+\cite{www-rivanna}
 \TODO{description of Rivanna on web page is not accurate.}
 
-\subsection{Rutherfordlab ???}
+\subsection{Pearl}
 \label{id:ral}
 
-Rutherfordlab (RAL)
+Rutherford Appleton Laboratory (RAL), UK 
+Pearl \cite{www-pearl-1}
 
@@
 
-Oak Ridge National Labratory (ORNL), Summit
+Oak Ridge National Labratory (ORNL), Summit \cite{www-summit}
 
-\subsection{NERSC ???}
+\subsection{Perlmutter}
 \label{id:nersc}
@@
 National Energy Research Scientific Computing Center (NERSC), ...
+\cite{www-perlmutter}
 
@@
 
-Argonne National Laboratory (ANL), ...
+Argonne National Laboratory (ANL), ... \cite{www-aurora}
 
@@
 
-San Diego Super Computing Center, ...
+San Diego Super Computing Center, ... \cite(www-expanse}
 
@@
     \centering
-    \begin{tabular}{|r|l|l|r|l|l|l|}
+    \begin{tabular}{|r|l|l|r|l|l|}
         \hline
-        Section & Organization & Machine  & Processors & GPUs & GPU Type  & Commissioned \\ 
+        Section & Organization & Machine  & Processors & GPUs &  Commissioned \\ 
         \hline
         \hline
-         \ref{id:uva}   & UVA       & Rivanna   &            &      &           & \\
-         \ref{id:ral}   & RAL       &           &            &      &           & \\
-         \ref{id:ornl}  & ORNL      &  Summit   &            &      &           & \\
-         \ref{id:nersc} & NERSC     &           &            &      &           & \\
-         \ref{id:anl}   & ANL       &           &            &      &           & \\
-         \ref{id:sdsc}  & SDSC      &           &            &      &           & \\
-         \ref{id:dgx}   & DGX Staion A100   &           & 1          &  AMD EPYC 7742 64-Core Processor   & A100 80GB  & Fox will know\\
+         \ref{id:anl}   & ANL       &  Aurora \cite{www-aurora}         &            &      &    ??? 2022       \\
+         \ref{id:ornl}  & ORNL      &  Summit \cite{www-summit}  &  27000  & NVIDIA Volta GPUs    &            \\
+         \ref{id:nersc} & NERSC     &  Perlmutter \cite{www-perlmutter}         &    6000    & NVIDIA A100 & Jan 2022        \\
+         \ref{id:sdsc}  & SDSC      &  Expanse \cite{www-expanse}         & 200&     NVIDIA V100 SMX2     &            \\
+         \ref{id:uva}   & UVA       & Rivanna \cite{www-rivanna}  &      8       & NVIDIA A100 80GB     & Feb 2022 \\
+         \ref{id:ral}   & RAL       &  Pearl \cite{www-pearl-1}        &   16         &  NVIDIA V100   &            \\
+         \ref{id:dgx}   & Fox & DGX Staion A100 \cite{www-dgx-station-a100}  & 4 &  NVIDIA A100 80GB  & May 2021\\
          \hline
@@
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{references,sample-base}
+\bibliography{references}
 
-\clearpage
+\section*{Biographies}
+
+\TODO{wrapfigure does somehow not work in this template}
+
+{\bf Thomas Butler}
+\TODO{Thomas: fill out. See Gregor's example}
+
+{\bf Robert Knuuti}
+is a Graduate Student at University of Virginia's School of Data Science.
+He has over 10 years experience in system architecture and software engineering, and specializes in Development Operations and Cloud Computing.
+\TODO{Not sure what to add about research organizations; I'm generally not affiliated.}  He has constructed air gapped Continuous Integration and Continuous Delivery systems for multiple organizations each supporting more than 100 developers and has facilitated the construction of repeatable, tractable builds for users of these systems.
+
+{\bf Jake Kolessar}
+\TODO{Jake: fill out. See Gregor's example}
+
+{\bf Geoffrey C. Fox}
+\TODO{Geoffery: fill out}
+
+%\begin{wrapfigure}{r}{0.25\columnwidth}
+%\includegraphics[width=0.24\columnwidth]{images/bio/gregor.png}
+%\end{wrapfigure}
+
+{\bf Gregor von Laszewski}
+is a Research Professor at University of Virginia.
+He has more than 30 years of experience in parallel and distributed computing.
+Selected research organizations he worked in include NASA, Argonne National Laboratory, Indiana University.
+He is proud to have been contributing members of researchteams that invented hybrid parallel genetic algorithms, Grid computing, and the first large scale academic hybrid cloud, as well as the larges scientific bibliometric analysis of XSEDE in the world.
 
@@
 \end{itemize}
+
+\clearpage
 
@@
 Additionally, it might be good to decide on some core roles to make sure things don't fall through the cracks (roles like recorder, scheduler, reviewer, et al...); the non-technical stuff for now, and then we can get into the more specific things once we pass the initial knowledge rampup}
+\TODO{We now have a project board for this semester, milestones, and issues.}
 

### 2022-02-19 paper-capstone-earthquake.tex
table addons for hardware links

--- 2022-01-28 paper-capstone-earthquake.tex
@@
-\documentclass[sigplan,screen]{acmart}
+\documentclass[sigplan,screen]{format/acmart}
 \usepackage{todonotes}
+\usepackage{graphicx}
+\usepackage{wrapfig}
+\usepackage[pdf]{graphviz}
+\usepackage{xpatch}
+\usepackage{adjustbox}
+\usepackage{makecell} 
+\usepackage{ifthen}
 
 \newcommand{\TODO}[1]{\todo[inline]{#1}}
+%\providecommand{\mlperf}{MLPerf\texttrademark{}}
+%\providecommand{\mlcube}{MLCube\texttrademark{}}
+\providecommand{\mlperf}{MLPerf}
+\providecommand{\mlcube}{MLCube}
 
-\setcopyright{acmcopyright}
-\copyrightyear{2022}
-\acmYear{2022}
-\acmDOI{XXXXXXX.XXXXXXX}
+%% Not sure what's the correct copyright to set in this template...  See the below link for
+%% the allowed options in Table 3.  Would this be acmlicensed as we're retaining copyright,
+%% but are allowing ACM's usage?  Or are we publishing elsewhere and this should be none?
+%% https://www.acm.org/binaries/content/assets/publications/consolidated-tex-template/acmart.pdf#page=18
+\setcopyright{none}
+%\setcopyright{Apache 2.0}
+%\setcopyright{rightsretained}
+\copyrightyear{}
+%\acmYear{2022}
+\acmDOI{}
 
-\acmConference[Conference acronym 'XX]{Make sure to enter the correct
-  conference title from your rights confirmation emai}{June 03--05,
-  2018}{Woodstock, NY}
-%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%  June 03--05, 2018, Woodstock, NY} 
-\acmPrice{15.00}
-\acmISBN{978-1-4503-XXXX-X/18/06}
+\acmConference[T.
+Buttler, R.
+Knuuti, J, Kolesar, Geoffrey C.
+Fox, Gregor von Laszewski, J.
+Fox, University of Virginia Data Science Capstone 2022]{MLCommons Earthquake Science Benchmark}{Jan.
+22 -- Aug.
+31 2022}{Charlotsville, VA, US}
+\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
+  June 03--05, 2018, Woodstock, NY} 
+\acmPrice{}
+\acmISBN{}
 %%\acmSubmissionID{123-A56-BU3}
@@
 
+\makeatletter
+\newcommand*{\addFileDependency}[1]{% argument=file name and extension
+  \typeout{(#1)}
+  \@addtofilelist{#1}
+  \IfFileExists{#1}{}{\typeout{No file #1.}}
+}
+\makeatother
+\xpretocmd{\digraph}{\addFileDependency{#2.dot}}{}{}
+% Removes margin issue
+% TODO - remove this line when we're ready to submit
+\setlength{\marginparwidth}{2cm}
 \begin{document}
@@
 \author{Thomas Butler}
-\email{trovato@corporation.com}
-\author{G.K.M. Tobin}
-\email{webmaster@marysville-ohio.com}
+\orcid{0000-0003-1515-0098}
 \affiliation{%
@@
   \country{USA}}
+ \email{Tommysbutler@gmail.com}
 
 \author{Robert Knuuti}
+\orcid{0000-0003-2614-0899}
 \affiliation{%
@@
 
-\renewcommand{\shortauthors}{Buttler, Knuuti, Kolesar, Fox, von Laszewski}
+\author{Judy Fox}
+\affiliation{%
+  \institution{University of Virginia}
+  \streetaddress{TBD}
+  \city{Charlotsville, VA}
+  \country{USA}}
+\email{cwk9mp@virginia.edu}
+
+\renewcommand{\shortauthors}{Butler, Knuuti, Kolesar, Fox, von Laszewski}
 
@@
 
+Several research efforts have been executed over the years to establish a complex set of factors that accurately predict where an earthquake could occur, and what impacts an earthquake would cause to the environment when they occur.
+Not only is this an algorithmically complex problem, but it is also computationally dense and requires high performance hardware to obtain these predictions\cite{fox2022aiforscience}.
+The intersection of these two elements create an ideal environment for building fair benchmarks.
+By establishing a rigorous process for constructing and running deep learning models the basis for comparing not just computational power in the High Performance Computing ecosystem, but also establish a benchmarks of new scientific discoveries from its execution.
+
+For this study, we review the following model architectures across various hardware platforms to establish this benchmark:
+\begin{itemize}
+    \item Long Short-Term Memory (LSTM) Recurrent Neural Network (RNN)
+    \item Spacio-Temporal Transformer Network (S2TN) as a Science Transformer
+    \item Temporal Fusion Transformer (TFT)
+\end{itemize}
+
+All three models leverage Deep Learning algorithms based upon transformer models.
+For each model, the benchmark analysis across the following deep learning frameworks:
+
+\begin{itemize}
+    \item Tensorflow (original reference)
+    \item Keras (original reference)
+    \item MXNet
+    \item Pytorch.
+\end{itemize}
+
+Each framework have their own performance measures that will shape the reference benchmark implementation.
+
 \subsection{MLCommons}
+
+MLCommons is an open engineering organization that seeks to unify the engineering and machine learning communities through system benchmarks, open data, and best practices \cite{www-mlcommons}.
+By using open data, a reference implementation for benchmarking, and executing best practices, the necessary rigor enables the construction repeatable research to establish a basis when performing future analysis.
+Using MLCommons organizational standards set by \mlperf{}, the dimensions of a the minimally viable solution for benchmarking are know and to establish our comparison criteria across our models and the executions on the different hardware sets\cite{mattson2019mlperf}.
+
+\TODO{Need to confirm what the criteria of the benchmark submission is.}
+
+\TODO{Address measure of Scientific Discovery, which is unique to the science benchmarks.}
+
+\subsection{Why Build a Science Benchmark?}
+
+The \mlperf{} framework already have several feature complete benchmarks that can judge system performance, however these benchmarks are centered on well known algorithms with solutions that are well known.
+Additionally, some CPU architectures create optimized instructions to solve common algorithmic problems at the microcode level, creating a performance boost when these prepared instructions are encountered\cite{cheong1997optimizer}.
+While these measures are still of great importance, they do not reflect the performance of systems based on non-deterministic measures that exist in nature.
+This makes benchmarking using a scientific dataset an ideal candidate for calculating performance, as there's an inherit randomness that's difficult to "game" this system and build predictions.
+This is especially true when attempting to predict the next occurrence of an earthquake, as complex models are required to gain predictive insights\cite{www-murakami2021}.
+
+\subsection{What is MLCommons Science Research?}
+
+MLCommons Science Research group is involved in edge and data center issues, end-to-end systems, inference and training.
+MLCommons Science Research focuses on Science applications and not industry.
+There are similarities between industry and science, for example image data is similar but some algorithms can be very different.
+The best example of this is particle physics experiments, quarks and lumens are not a common thing to model in industry \cite{www-mlcommons-science}.
+
+The primary goal of MLCommons Science Research is to develop optimal methods to improve both speed and efficiency of scientific outcomes.
+This is done by measure how different models perform on different scientific datasets, in particular score the models based on how well they do on scientific discovery.
+This could be anywhere from accuracy of a model to time to reach  specific accuracy target when new data is introduced.
+In general, how well each model matches the goals of that particular scientific field.
+By using verified datasets following best standardized practice on how to build and exchange models using the reference materials on benchmarked machines, knowledge can flow more easily and freely between various researchers.
+By having these standard methods, both the monetary cost and time can be reduced between researcher regardless of their speciality.
+
+Currently MLCommons Science Research has four datasets based on climate, material science, medicine and earthquakes with a fifth dataset on plasma physics being added soon.
+Each dataset has their own method used to solve their scientific problem \cite{www-mlcommons-science}.
+
+\subsection{What is \mlcube{}?}
+
+\mlcube{} is  a set of standards and software developed by MLCommons for delivering machine learning models across container runtimes such as docker, kubernetes, singularity, or for automating the model's execution over ssh.
+\mlcube{} provides a common configuration format and interface such that consumers of a model do not need to worry about the underlying configuration and baseline of supporting libraries, and enables the user to focus on the model itself\cite{www-mlcube}.
+
+There are two targets when building \mlcube{} models:
+
+\begin{itemize}
+    \item \textbf{Developer.}  The developer produces a YAML configuration file that describes the target execution details including platform, runtime dependencies, inputs, outputs, metadata about the model, and the source code.
+Additionally, this configuration file offers branching paths based on what \mlcube{} runtime backend is selected (for example, the container image to use for docker can be configured differently than a kubeflow configuration).
+This is so the implementation details can be explicitly addressed, if necessary, by the model developer and abstracts this complexity away from consumers.
+    \item \textbf{User.}  The user of the system will receive the \mlcube{} code repository and execute a command that identifies which platform they intend to run the model on and any type of input parameters to the model detailed by the developer.
+\end{itemize}
+
+As described by the \mlcube{} development team, this ecosystem establishes a contract between model developers and consumers of that model so that results can be easily shared and provide a framework to produce consistent results.
+
+The MLCommons Earthquake Science Benchmark models target the \mlcube{} mechanism for delivery.
 
@@
 
+Earthquake prediction is a complex and old problem.
+There are two major tasks related to earthquakes.
+One is forecasting earthquake occurrences/likelihoods.
+Another is to predict damage once an earthquake does happen.
+The first is difficult because the details of the underground plates and the friction laws between them are not known and furthermore the earthquake causes phase transitions between plates which makes for unpredictable movement.
+The second, we need to predict earthquake energy wave movements.
+Since both of these problems are so complex with many hidden variables and the equations governing the phenomenon are unknown or incomplete, deep learning can be use to learn the various hidden patterns in the data which the model can then use to forecast earthquakes into the future.
+\cite{fox2021earthquake}
+
+\subsection{Energy-based averaging, the log loss equation}
+
+\TODO{This section is in progress! Enter Energy Weighted Quantities equations picture below}
+
+This function maximizes the largest earthquake energy values for the 2 week period. 
+Magnitude averaged over the bin
+All input values and predictions are independently normalized and have a maximum modulus of 1 over all space and time values. \cite{fox2022aiforscience}
+
+\subsection{Earthquake Data turned into Spatial bag Time Series}
+
+\TODO{This section is in progress!}
+
+The data being used is the Southern California location data, between 32 to 36 lat -120 to -114 long.
+The data is recorded between 1950-2020.
+Each location point is is 0.1 by 0.1 degrees or ll km by ll km in size or 2400 points in total.
+The points are then put into spatial bag which are based on a bunch of different location points, necessarily nearby spatially.
+Each location point has a time series associated with it.
+The time series data holds magnitude, depth, spatial position, and time.
+\cite{fox2021earthquake}
+
+The data was accumulated into 2 week samples using energy-based averaging, the log loss equation.
+
+\TODO{provide introduction to earthquake}
+
+\TODO{Review how these papers contribute to introduction}
+
+\begin{itemize}
+    \item Review IEEE Presentation \cite{fox2022aiforscience}
+    \item Review Fox paper \cite{fox2021earthquake}
+\end{itemize}
+
+\section{Timeseries Earthquake Prediction}
+
+\subsection{LSTM}
+
+\begin{figure}[htb]
+\resizebox{0.75\columnwidth}{!}{
+    \digraph{lstmarchitecture}{
+        node [shape=box style=rounded color=grey];
+        Embedder -> "2-Layer LSTM" -> "Output Mapper";
+    }
+}
+\caption{LSTM Model}\label{fig:lstm-architecture}
+\end{figure}
+
+\subsection{TFT}
+
+\TODO{Fix this figure}
+
+\begin{figure}[htb]
+\resizebox{0.75\columnwidth}{!}{
+%    \digraph{tftarchitecture}{
+%        compound=true;
+%        newrank=true;
+%        rankdir="LR";
+%        node [shape=box style=rounded color=grey];
+
+%        start[label="Embedders"];
+%        ascbackward[label="2-layer LSTM as Backward encoder"];
+%        ascforward[label="2-layer LSTM as Forward decoder"];
+%        ta[label="Temporal Attention"];
+
+ %       start -> ta [lhead="clusterstaticcontext", constraint=false];
+ %       ta -> "Output Mappers";
+ %       
+ %       subgraph clusterstaticcontext {
+ %           rankdir="TB";
+ %           label="Static Context";
+ %           {rank="same"; ascbackward; ascforward;}
+ %           ascbackward -> ta;
+ %           ascbackward -> ascforward;
+ %           ascforward -> ta;
+%        }
+%    }
+}
+\caption{TFT Model}\label{fig:tft-architecture}
+\end{figure}
+
+%% Toying with tikz for generating the grpahic.
+%\tikz [nodes={draw}] {
+%    \node (a) {Embedders};
+%    \node (b) [below left of=a] {2-layer LSTM as Backward Encoder};
+%    \node (c) [below right of=a] {2-layer LSTM as Forward Decoder};
+%    \node (d) [below right of=b] {Temporal Attention};
+%    \node (e) [below of=d] {Output Mappers};
+%    \draw[->] (a) -- (b);
+%    \draw[->] (a) -- (c);
+%    \draw[->] (b) -- (c);
+%    \draw[->] (b) -- (d);
+%    \draw[->] (c) -- (d);
+%    \draw[->] (d) -- (e);
+%}
+
+%\begin{figure}[h]
+%  \centering
+%  \includegraphics[width=\linewidth]{sample-franklin}
+%  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
+%    Ewing, Inc. [Public domain], via Wikimedia
+%    Commons. (\url{https://goo.gl/VLCRBB}).}
+%  \Description{A woman and a girl in white dresses sit in an open car.}
+% \end{figure}
+
+\section{Benchmarks}
+
+\subsection{Gregor's MNIST}
+
+Gregors MNIST benchmark results are summarized in Table~\ref{tab:mnist}.
+
+\begin{table*}[!ht]
+\caption{Gregors MNIST Benchmarks}\label{tab:mnist}
+    \centering
+      \begin{adjustbox}{max width=\textwidth}
+    \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
+    \hline
+        Machine & real & user & sys & Driver & CUDA & GPU & \makecell{Date CPU \\released} & CPU & \makecell{Date CPU \\released} \\ \hline
+        Gregors Machine & 0m11.534s & 0m13.914s & 0m05.186s & 510.47.03 & 11.6 & Gigabyte RTX3070 TI & May 31, 2021 & AMD 5950X & Nov 2020 \\ \hline
+        Fox DGX & 0m19.987s & 5m12.991s & 0m49.266s & 450.142.00 & 11.0 & NVIDIA A100 80GB & & AMD EPYC 7742 64-Core & Aug 2019 \\ \hline
+        Rivanna A100 & 0m29.263s & 0m14.585s & 0m7.399ss & 470.82.01 & 11.4 & NVIDIA A100-SXM4-40GB & May 14, 2020 & Intel(R) Xeon(R) CPU E5-2630 v3 @ 2.40GHz & Q3  2014 \\ \hline
+        MacBook Pro & 0m31.01s & 0m25.26s & 130\% & N/A & N/A & N/A & N/A & M1 Max 66GB & Nov 2021 \\ \hline
+        Rivanna P100 & 0m35.732s & 0m17.253s & 0m7.595s & 470.82.01 & 11.4 & Tesla P100-PCIE & & Intel(R) Xeon(R) CPU E5-2630 v3 @ 2.40GHz & Q3  2014 \\ \hline
+        Rivanna V100 & 0m43.160s & 0m15.510s & 0m6.894s & 470.82.01 & 11.4 & Tesla V100-SXM2 & & Intel(R) Xeon(R) CPU E5-2630 v3 @ 2.40GHz & Q3  2014 \\ \hline
+        Rivanna K80 & 0m57.588s & 0m20.322s & 0m9.612s & 470.82.01 & 11.4 & NVIDIA TESLA K80 & & Intel(R) Xeon(R) CPU E5-2630 v3 @ 2.40GHz & Q3  2014 \\ \hline
+        Rivanna Frontend & 1m11.535s & 1m00.780s & 0m10.352s & N/A & N/A & N/A & & Intel(R) Xeon(R) CPU E5-2630 v3 @ 2.40GHz & Q3  2014 \\ \hline
+    \end{tabular}
+    \end{adjustbox}
+\end{table*}
+
+\subsection{Hardware}
+
+\subsection{Aurora}
+\label{id:anl}
+
+The Aurora supercomputer, sponsored by the U.S.
+Department of Energy (DOE), is part of the Argonne Leadership Computing Facility (ACLF).
+It was developed with the help of industry experts from Intel and Cray.
+Aurora offers HPC for data analytics and AI development with a peak performance of over 2 Exaflop DP.
+The system is based on the HPE Cray EX supercomputer platform and offers over 9,000 compute nodes.
+
+Argonne National Laboratory (ANL), ... \cite{www-aurora}
+
+\subsection{Summit}
+\label{id:ornl}
+
+The Summit supercomputer, sponsored by the U.S.
+Department of Energy (DOE), is located in the Oak Ridge Leadership Computing Facility (OLCF).
+The open science computing system is available to government, academic, and industry researchers to solve complex problems in all fields of science.
+The system is powered by 4,608 nodes which each house multiple IBM POWER9 CPUs and NVIDIA Volta GPUs on top of half a terabyte of coherent memory.
+
+Oak Ridge National Laboratory (ORNL), Summit \cite{www-summit}
+
+\subsection{Perlmutter}
+\label{id:nersc}
+
+The Perlmutter supercomputer, sponsored by the U.S.
+Department of Energy (DOE), is located in Shyh Wang Hall at Berkeley Lab and is operated by the National Energy Research Scientific Computing Center (NERSC).
+The supercomputer has a 12 cabinet GPU systems with over 1,500 nodes and a 12 cabinet CPU system containing over 3,000 nodes.
+It is supported by 35 petabytes of all-flash storage which can move data at 5 terabytes/sec.
+
+National Energy Research Scientific Computing Center (NERSC), ...
+\cite{www-perlmutter}
+
+\subsection{SDSC}
+\label{id:sdsc} 
+\TODO{Which of the SDSC systems will we be benchmarking?}
+San Diego Super Computing Center, ... \cite{www-expanse}
+
+\subsection{University of Virginia Rivanna}
+\label{id:uva}
+
+UVA, Rivanna, A100, RTX3090, others may not have enough memory. 
+\cite{www-rivanna}
+\TODO{description of Rivanna on web page is not accurate.}
+
+Containers are available on Rivanna through Singularity \cite{www-singularity} that can be loaded with modules \cite{www-modules}.
+We plan to utilize singularity, but target the formulation of the container in docker and than transform it to singularity.
+As singularity is installed also on other machines, we will leverage this deployment also on them.
+
+\begin{figure}[htb]
+\resizebox{0.75\columnwidth}{!}{
+\digraph{singularity}{
+    node [shape=box style=rounded color=grey];
+    Application -> GitHub
+    "Dockerfile" -> "Docker Container" -> "Singularity Conversion" -> "Singularity Container" -> "Batch job (SLURM)"
+}
+}
+\caption{Containerized Science Applications in Singularity}
+\end{figure}
+
+\subsection{Pearl}
+\label{id:ral}
+
+The PEARL system, funded by the Alan Turing Institute, is hosted at the Rutherford Appleton Laboratory by the Science and Technology Facilities Council (STFC).
+It was developed for the advancement of AI and Machine Learning research with large-scale scientific datasets, already contributing to the fields of materials, environmental and life sciences, and astronomy.
+It is comprised of two NVIDIA DGX2 nodes, 32 GPUs, 3TB of system memory, 1TB of GPU memory, and 600TB of high-speed storage.
+
+Rutherford Appleton Laboratory (RAL), UK 
+Pearl \cite{www-pearl-1}
+
+\subsection {Nvidia DGX Workstation}
+\label{id:dgx}
+
+The NVIDIA DGX Station A100 is a commercially available product that can provide powerful computing resources to all industries.
+The DGX Station is an AI resource that can run parallel jobs and support training, inference, and data analytics.
+It contains 4 NVIDIA A100 GPUs and up to 320 GBs of GPU memory.
+With the use of Multi-Instance GPU (MIG), each A100 GPU can be partitioned into 7 isolated instances.
+
+DGX Fox \cite{www-dgx-station-a100}
+
+\newcommand{\nvidia}[1]{
+    \ifthenelse{\equal{#1}{A100}}{\href{https://www.nvidia.com/en-us/data-center/a100/}{NVIDIA #1}}{} 
+    \ifthenelse{\equal{#1}{V100}}{\href{https://www.nvidia.com/en-us/data-center/a100/}{NVIDIA #1}}{}
+    \ifthenelse{\equal{#1}{K80}}{\href{https://www.nvidia.com/en-gb/data-center/tesla-k80/}{NVIDIA #1}}{}
+    \ifthenelse{\equal{#1}{P100}}{\href{https://www.nvidia.com/en-us/data-center/tesla-p100/}{NVIDIA #1}}{}
+    \ifthenelse{\equal{#1}{RTX3090}}{\href{https://www.nvidia.com/en-us/geforce/graphics-cards/30-series/rtx-3090/}{NVIDIA #1}}{}
+    \ifthenelse{\equal{#1}{RTX2080TI}}{\href{https://www.nvidia.com/en-us/data-center/tesla-p100/}{NVIDIA #1}}{}
+
+%\newcommand{\nvidiaV}{\href{https://www.nvidia.com/en-us/data-center/v100/}{NVIDIA V100}}
+
+\begin{table*}[htb]
+    \caption{Overview of compute resources.}
+    \label{tab:my_label}
+    \centering
+%\begin{adjustbox}{angle=90}
+\begin{adjustbox}{max width=\textwidth}
+    \begin{tabular}{|r|l|ll|r|l|l|l|l|l|}
+        \hline
+        Section & Organization   & Machine                         & & Processors    & GPUs             & \makecell{Memory\\/Device} &  \makecell{GPU\\/Device} & \makecell{No.
+of\\ nodes} & Commissioned \\
+        \hline
+        \hline
+         \ref{id:anl}   & ANL    & Aurora & \cite{www-aurora}         &              &                  & & & &    ??? 2022   \\ \hline
+         \ref{id:ornl}  & ORNL   & Summit &\cite{www-summit}         &        27000 & NVIDIA Volta     & & & &               \\ \hline
+         \ref{id:nersc} & NERSC  & Perlmutter & \cite{www-perlmutter} &         6000 & \nvidia{A100}      & & & & Jan 2022      \\ \hline
+         \ref{id:sdsc}  & SDSC   & Expanse & \cite{www-expanse}       &          200 & NVIDIA V100 SMX2 & & & &               \\ \hline
+         \ref{id:uva}   & UVA    & Rivanna &\cite{www-rivanna}       &            8 & \nvidia{A100}      & 80GB & ? & &  Feb 2022      \\
+            &    &  &       &            8 & \nvidia{A100}      & 40GB & 8 & &       \\
+            &    &  &       &            ? & \nvidid{RTX3090}   & 24GB & ? & ? & 2021          \\
+            &    &  &       &            ? & \nvidia{K80}       & 11GB & 8 & 9 & 2021          \\
+            &    &  &       &            ? & \nvidia{V100}      & 16GB & 4 & 1 & 2021          \\
+            &    &  &       &            ? & \nvidia{V100}      & 32GB & 4 & 12 & 2021          \\
+            &    &  &       &            ? & \nvidia{P100}      & 12GB & 4 & 3 & 2021          \\
+            &    &  &       &            ? & \nvidia{RTX2080TI} & 11GB & 10 & 2 & 2021          \\
+         \hline
+         \ref{id:ral}   & RAL    & Pearl &\cite{www-pearl-1}         &           16 & \nvidiaV      & & & &  \\ \hline
+         \ref{id:dgx}   & Fox    & DGX Station A100 &\cite{www-dgx-station-a100} & 4 & \nvidiaA      & 80GB & 4 & 1 & May 2021      \\
+         \hline
+    \end{tabular}
+    \end{adjustbox}
+\end{table*}
+
+Rivanna: \url{https://www.rc.virginia.edu/userinfo/rivanna/overview/}
+
+\TODO{Add your personal computers to this table \url{https://github.com/Data-ScienceHub/mlcommons-science/issues/37}}
+\begin{table*}[htb]
+    \caption{Overview of compute resources.}
+    \label{tab:my_label}
+    \centering
+%\begin{adjustbox}{angle=90}
+\begin{adjustbox}{max width=\textwidth}
+    \begin{tabular}{|r|l|l|r|l|l|l|l|l|}
+        \hline
+        Section & Organization   & Machine                         & Processors    & GPUs             & \makecell{Memory\\/Device} &  \makecell{GPU\\/Device} & \makecell{No.
+of\\ nodes} & Commissioned \\
+        \hline
+        \hline
+         \ref{id:gcomputer}   & 
+         Gregor    & 
+         5950X   &  
+         1 &   
+         \href{https://www.gigabyte.com/Graphics-Card/GV-N307TGAMING-OC-8GD#kf}{GIGABYTE Gaming RTX 3070TI}  & 
+         8GB & 
+         1 & 
+         1 & 
+         Sep 2021   \\
+            & 
+            & 
+            &  
+            &   
+         \href{https://rog.asus.com/us/graphics-cards/graphics-cards/rog-strix/rog-strix-rtx3090-o24g-gaming-model/}{ASUS ROG Strix RTX 3090} & 
+         24GB & 
+         1 & 
+         1 & 
+         Feb 2022   \\
+         \hline
+         \ref{id:tcomputer}   & Thomas.A  & i7-8750H &  1 &   NVIDIA GeForce RTX 2070 with Max-Q Design  & 16GB & 1 & 1 & 2019   \\
+         \hline
+         \ref{id:t2computer}   & Thomas.B  & i7-7700HQ &  1 &   NVIDIA GeForce GTX 1060  & 16GB & 1 & 1 & 2017   \\
+         \hline
+    \end{tabular}
+    \end{adjustbox}
+\end{table*}
+
+\begin{acks}
+\TODO{Geoffrey: please add acknowledgement for funding, other acks}
+\end{acks}
+
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{references}
+
+\section*{Biographies}
+
+\TODO{wrapfigure does somehow not work in this template}
+
+{\bf Thomas Butler}
+is a Graduate Student at University of Virginia's School of Data Science.
+His undergraduate degree is in Biomedical engineering.
+He has over 8 years of experience in the Infertility field helping patients, jointly running a Andrology lab, and contributing research to advance the field through joint research on how AMH effects pregnancy outcomes and sperm antibodies effect PSA.
+He has a certificate in Data Analytics from Georgia Institute of Technology.
+
+{\bf Robert Knuuti}
+is a Graduate Student at University of Virginia's School of Data Science.
+He has over 10 years experience in system architecture and software engineering, and specializes in Development Operations and Cloud Computing.
+He has constructed air gapped Continuous Integration and Continuous Delivery systems for multiple organizations each supporting more than 100 developers and has facilitated the construction of repeatable, tractable builds for users of these systems.
+
+{\bf Jake Kolessar}
+{is a Graduate Student at the University of Virginia's School of Data Science.
+He has a background in mechanical engineering and 2 years of experience as a Modeling, Simulation & Analysis Engineer.
+He has supported the software design and development of modeling capabilities for event simulation products as well as the integration of models into the simulation framework.}
+
+{\bf Geoffrey C. Fox}
+\TODO{Geoffery: fill out}
+
+%\begin{wrapfigure}{r}{0.25\columnwidth}
+%\includegraphics[width=0.24\columnwidth]{images/bio/gregor.png}
+%\end{wrapfigure}
+
+{\bf Gregor von Laszewski}
+is a Research Professor at University of Virginia.
+He has more than 30 years of experience in parallel and distributed computing.
+Selected research organizations he worked in include NASA, Argonne National Laboratory, Indiana University.
+He is proud to have been contributing members of research teams that invented hybrid parallel genetic algorithms, Grid computing, and the first large scale academic hybrid cloud, as well as the larges scientific bibliometric analysis of XSEDE in the world.
+
+\appendix
+
+\section{Manuals Developed}
+
+The following manuals have been developed by the team:
+
+\begin{itemize}
+    \item Introduction to Python \cite{las-intro-python}
+\end{itemize}
+
+\clearpage
+
+\section{Todo}
+
+\subsection{Manual List}
+\section{TODO and integrate}
+
 Things to do:
+
 \begin{itemize}
-    %\item add presentation as cite,
-    \item Review IEEE Presentation \cite{fox2022aiforscience}
+    %% \item add presentation as cite,
     \item Review MLCommons \cite{www-mlcommons} 
     \item Use jabref \cite{www-jabrefg-org} for citation management
-    \item Review Fox paper \cite{fox2021earthquake}
     \item Frequently check github \cite{www-mlcommons-eathquake}
@@
     \item Learn how to selfdeploy jupyterlab \cite{www-jupyterlab}
-    \TODO{Is this the same as jupytext?}
-
     \item Begin to learn about papermill \cite{www-papermill} \TODO{Rivanna's core configurations might be a bit limited on this front.
@@
     \TODO{MLCommons project that can target a few platforms using a YAML contract.  Once we solve the target environment, we can likely target porting this way.}
-    \TODO{Also, good news is Rivanna has a Singularity module  as an lmod (3.7.1), so we might be able to use that as a targeted runner without having to setup too much additional scaffolding.
-    And if things work right, we could use the k8s module to do local deployments if we wanted to test out alternate execution paths. }
-    \item mlcube \cite{www-mlcube} 
+
+    \item \mlcube{} \cite{www-mlcube} 
 \end{itemize}
 
-%\begin{figure}[h]
-%  \centering
-%  \includegraphics[width=\linewidth]{sample-franklin}
-%  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
-%    Ewing, Inc. [Public domain], via Wikimedia
-%    Commons. (\url{https://goo.gl/VLCRBB}).}
-%  \Description{A woman and a girl in white dresses sit in an open car.}
-% \end{figure}
+\subsection{Generated List}
+\listoftodos{}
 
-\section{Benchmarks}
-
-\subsection{Hardware}
-
-\subsection{University of Virginia Rivanna}
-\label{id:uva}
-
-UVA, Rivanna, A100, RTX3090, others may not have enough memory. 
-\TODO{description of Rivanna on web page is not accurate.}
-
-\subsection{Rutherfordlab ???}
-\label{id:ral}
-
-Rutherfordlab (RAL)
-
-\subsection{Summit}
-\label{id:ornl}
-
-Oak Ridge National Labratory (ORNL), Summit
-
-\subsection{NERSC ???}
-\label{id:nersc}
-
-National Energy Research Scientific Computing Center (NERSC), ...
-
-\subsection{Aurora}
-\label{id:anl}
-
-Argonne National Laboratory (ANL), ...
-
-\subsection{SDSC}
-\label{id:sdcs}
-
-San Diego Super Computing Center, ...
-
-\subsection {Nvidia DGX Workstation}
-
-\label{id:dgx} DGX Fox \cite{www-dgx-station-a100}
-
-\begin{table*}[htb]
-    \caption{Overview of compute resources.}
-    \label{tab:my_label}
-    \centering
-    \begin{tabular}{|r|l|l|r|l|l|l|}
-        \hline
-        Section & Organization & Machine  & Processors & GPUs & GPU Type  & Commissioned \\ 
-        \hline
-        \hline
-         \ref{id:uva}   & UVA       & Rivanna   &            &      &           & \\
-         \ref{id:ral}   & RAL       &           &            &      &           & \\
-         \ref{id:ornl}  & ORNL      &  Summit   &            &      &           & \\
-         \ref{id:nersc} & NERSC     &           &            &      &           & \\
-         \ref{id:anl}   & ANL       &           &            &      &           & \\
-         \ref{id:sdsc}  & SDSC      &           &            &      &           & \\
-         \ref{id:dgx}   & DGX Staion A100   &           & 1          &  AMD EPYC 7742 64-Core Processor   & A100 80GB  & Fox will know\\
-         \hline
-    \end{tabular}
-\end{table*}
-
-\begin{acks}
-\TODO{Geoffrey: please add acknowledgement for funding, other acks}
-\end{acks}
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{references,sample-base}
-
-\clearpage
-
-\appendix
-
-\section{Manuals Developed}
-
-The following manuals have been developed by the team:
+To find a general list of todo actions, consult
 
 \begin{itemize}
-    \item Introduction to Python \cite{las-intro-python}
+\item The \href{https://github.com/cybertraining-dsc/capstone-eartquake/blob/main/TODO.md}{TODO.md} 
+\item The \href{https://github.com/Data-ScienceHub/mlcommons-science/projects/1}{GitHub Project}
 \end{itemize}
 
-\section{Todo}
+\section{Progress Reports}
 
-\listoftodos
+\subsection{Progress Report A}
 
-Robert: I am moving the rough notes into our github repo and made a
+Dates covered: Jan 27 - Feb 23rd
+
+Due: Feb 24th, 2022
 
 \begin{itemize}
-\item \href{https://github.com/cybertraining-dsc/capstone-eartquake/blob/main/TODO.md}{TODO.md} 
+\item Reviewed background information \footnote{\url{https://github.com/Data-ScienceHub/mlcommons-science/issues/1}}
+    \begin{itemize}
+    \item Earthquake Nowcasting with Deep Learning
+    \item MLCommons Benchmark Presentation
+    \item Attention is ALl you Need
+    \item MLCommons TEvolOp Earthquake Forcasting objectives
+    \item Learning on Temporal Fusion Transformers
+    \item MLCommons and AI for Science illustrated by Deep Learning for Geospatial Time Series
+    \end{itemize}
+\item Reviewed background technologies \footnote{\url{https://github.com/Data-ScienceHub/mlcommons-science/issues/2}}
+    \begin{itemize}
+    \item lmod - lua module and environment management (used on rivanna)
+    \item pip / conda env - python package management
+    \item SLURM - an HPC batch job runner
+    \item papermill - parameterized jupyter notebooks
+    \item horovod - distributed framework for running multiple deep learning frameworks
+    \item \mlcube{} - MLCommons repeatable machine learning framework with multiple backends.
+    \item singularity - container runtimes targeting HPC clusters
+    \item docker
+\end{itemize}
+\item Wrote introduction to paper (in progress) \footnote{\url{https://www.overleaf.com/project/61f1e28f076f4111c3ba927e}}
+    \begin{itemize}
+    \item added references to references.bib in overleaf
+    \end{itemize}
+\item Revised Intro to Python guide for conda installations so it doesn't interfere with pyenv or other shell commands.\footnote{\url{https://github.com/cloudmesh-community/book/pull/574}} \footnote{\url{https://github.com/cloudmesh-community/book/pull/575}}
+\item Solved conda environment and requirements.txt for colab notebook
+\footnote{\url{https://github.com/Data-ScienceHub/mlcommons-science/issues/12}} \footnote{\url{https://github.com/laszewsk/mlcommons/pull/1}} \footnote{\url{https://github.com/laszewsk/mlcommons/pull/3}}
+\item build docker container image that can run python 3.9.7 and conda
+\item verified access to GPU allocations in Rivanna
 \end{itemize}
 
-file that we can use to build a simple list of things we want to do.
-I'll leave them in the overleaf for now, but I feel like that's not a good home for that type of information given that it's mean to be the paper to submit on the subject.
+\subsection{Progress Report B}
 
-\TODO{Not sure if we want to talk about building tasks / kanban boards to track the different parts of work we'll be doing each week.
-Additionally, it might be good to decide on some core roles to make sure things don't fall through the cracks (roles like recorder, scheduler, reviewer, et al...); the non-technical stuff for now, and then we can get into the more specific things once we pass the initial knowledge rampup}
+Dates covered: Feb 24th - Mar 23rd
+
+Due: Mar 24th, 2022
+
+\subsection{Presentation}
+
+Dates covered: Mar 24th - April 27th
+
+Due: April 28th, 2022
 

### 2022-02-19 paper-capstone-earthquake.tex
sync
capstone

@@
+\PassOptionsToPackage{hyphens}{url}
 \documentclass[sigplan,screen]{format/acmart}
@@
 \usepackage{ifthen}
+\usepackage[numbers]{natbib}‎
+
+\usepackage[colorlinks = true,
+            linkcolor = blue,
+            urlcolor  = blue,
+            citecolor = blue,
+            anchorcolor = blue]{hyperref}
 
@@
 
-\acmConference[T.
-Buttler, R.
-Knuuti, J, Kolesar, Geoffrey C.
-Fox, Gregor von Laszewski, J.
-Fox, University of Virginia Data Science Capstone 2022]{MLCommons Earthquake Science Benchmark}{Jan.
-22 -- Aug.
-31 2022}{Charlotsville, VA, US}
-\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-  June 03--05, 2018, Woodstock, NY} 
+%\acmConference[T. Buttler, R. Knuuti, J, Kolesar, G. C. Fox,
+%  G. von Laszewski, J. Fox, University of Virginia Data Science
+%  Capstone 2022]{MLCommons Earthquake Science Benchmark}{Jan. 22 --
+%  Aug. 31 2022}{Charlotsville, VA, USA}
+
+\acmConference[Data Science
+  Capstone, University of Virginia, 2022]{MLCommons Earthquake Science Benchmark}{Jan. 22 --
+  Aug. 31 2022}{Charlotsville, VA, USA}
+
+\acmBooktitle{Woodstock '18:
+  ACM Symposium on Neural Gaze Detection, June 03--05, 2018,
+  Woodstock, NY}
+
 \acmPrice{}
@@
 \xpretocmd{\digraph}{\addFileDependency{#2.dot}}{}{}
-% Removes margin issue
-% TODO - remove this line when we're ready to submit
-\setlength{\marginparwidth}{2cm}
+
 \begin{document}
@@
   \institution{University of Virginia}
-  \streetaddress{TBD}
-  \city{Charlotsville, VA}
+  \city{Charoletteville}
   \country{USA}}
@@
   \institution{University of Virginia}
-  \streetaddress{TBD}
   \city{Charoletteville}
@@
   \institution{University of Virginia}
-  \streetaddress{TBD}
-  \city{Charoletteville}
+  \city{Charlottesville}
+  \state{VA}
+  \postcode{22911}
   \country{United States}}
@@
   \institution{University of Virginia}
-  \streetaddress{TBD}
-  \city{Charlotsville, VA}
+  \streetaddress{Biocomplexity Institute
+Town Center Four
+994 Research Park Boulevard}
+  \city{Charlottesville}    
+  \state{VA}
+  \postcode{22911}
   \country{USA}}
@@
   \institution{University of Virginia}
-  \streetaddress{TBD}
-  \city{Charlotsville, VA}
+  \city{Charlottesville}      
+  \state{VA}
+  \postcode{22911}
   \country{USA}}
@@
 
-\renewcommand{\shortauthors}{Butler, Knuuti, Kolesar, Fox, von Laszewski}
+\renewcommand{\shortauthors}{Butler, Knuuti, Kolesar, G Fox, G von Laszewski}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% THIS SECTION MAY NOT BE COPIED INTO ANOTHER PAPER WITHOUT MAKING GREGOR VON LASZEWSKI COAUTHOR
+%
+%
 \newcommand{\nvidia}[1]{
-    \ifthenelse{\equal{#1}{A100}}{\href{https://www.nvidia.com/en-us/data-center/a100/}{NVIDIA #1}}{} 
-    \ifthenelse{\equal{#1}{V100}}{\href{https://www.nvidia.com/en-us/data-center/a100/}{NVIDIA #1}}{}
-    \ifthenelse{\equal{#1}{K80}}{\href{https://www.nvidia.com/en-gb/data-center/tesla-k80/}{NVIDIA #1}}{}
-    \ifthenelse{\equal{#1}{P100}}{\href{https://www.nvidia.com/en-us/data-center/tesla-p100/}{NVIDIA #1}}{}
-    \ifthenelse{\equal{#1}{RTX3090}}{\href{https://www.nvidia.com/en-us/geforce/graphics-cards/30-series/rtx-3090/}{NVIDIA #1}}{}
-    \ifthenelse{\equal{#1}{RTX2080TI}}{\href{https://www.nvidia.com/en-us/data-center/tesla-p100/}{NVIDIA #1}}{}
-
+\ifthenelse{\equal{#1}{A100}}{\href{https://www.nvidia.com/en-us/data-center/a100/}{NVIDIA #1}}{}%
+\ifthenelse{\equal{#1}{V100}}{\href{https://www.nvidia.com/en-us/data-center/a100/}{NVIDIA #1}}{}%
+\ifthenelse{\equal{#1}{K80}}{\href{https://www.nvidia.com/en-gb/data-center/tesla-k80/}{NVIDIA #1}}{}%
+\ifthenelse{\equal{#1}{P100}}{\href{https://www.nvidia.com/en-us/data-center/tesla-p100/}{NVIDIA #1}}{}%
+\ifthenelse{\equal{#1}{RTX3090}}{\href{https://www.nvidia.com/en-us/geforce/graphics-cards/30-series/rtx-3090/}{NVIDIA #1}}{}%
+\ifthenelse{\equal{#1}{RTX2080TI}}{\href{https://www.nvidia.com/en-eu/geforce/20-series/}{NVIDIA #1}}{}%
+\ifthenelse{\equal{#1}{Volta}}{\href{https://www.nvidia.com/en-us/data-center/volta-gpu-architecture/}{NVIDIA #1}}{}%
+}
 %\newcommand{\nvidiaV}{\href{https://www.nvidia.com/en-us/data-center/v100/}{NVIDIA V100}}
@@
          \ref{id:anl}   & ANL    & Aurora & \cite{www-aurora}         &              &                  & & & &    ??? 2022   \\ \hline
-         \ref{id:ornl}  & ORNL   & Summit &\cite{www-summit}         &        27000 & NVIDIA Volta     & & & &               \\ \hline
+         \ref{id:ornl}  & ORNL   & Summit &\cite{www-summit}         &        27000 & \nvidia{volta}     & & & &               \\ \hline
          \ref{id:nersc} & NERSC  & Perlmutter & \cite{www-perlmutter} &         6000 & \nvidia{A100}      & & & & Jan 2022      \\ \hline
-         \ref{id:sdsc}  & SDSC   & Expanse & \cite{www-expanse}       &          200 & NVIDIA V100 SMX2 & & & &               \\ \hline
+         \ref{id:sdsc}  & SDSC   & Expanse & \cite{www-expanse}       &          200 & \nvidid{V100}
+         SMX2 & & & &               \\ \hline
          \ref{id:uva}   & UVA    & Rivanna &\cite{www-rivanna}       &            8 & \nvidia{A100}      & 80GB & ? & &  Feb 2022      \\
@@
          \hline
-         \ref{id:ral}   & RAL    & Pearl &\cite{www-pearl-1}         &           16 & \nvidiaV      & & & &  \\ \hline
-         \ref{id:dgx}   & Fox    & DGX Station A100 &\cite{www-dgx-station-a100} & 4 & \nvidiaA      & 80GB & 4 & 1 & May 2021      \\
+         \ref{id:ral}   & RAL    & Pearl &\cite{www-pearl-1}         &           16 & \nvidia{V100}      & & & &  \\ \hline
+         \ref{id:dgx}   & Fox    & DGX Station A100 &\cite{www-dgx-station-a100} & 4 & \nvidia{A100}      & 80GB & 4 & 1 & May 2021      \\
          \hline
@@
 
+%
+% THIS SECTION MAY NOT BE COPIED INTO ANOTHER PAPER WITHOUT MAKING GREGOR VON LASZEWSKI COAUTHOR
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{references}
+\bibliography{paper-capstone-earthquake}
 
@@
 His undergraduate degree is in Biomedical engineering.
-He has over 8 years of experience in the Infertility field helping patients, jointly running a Andrology lab, and contributing research to advance the field through joint research on how AMH effects pregnancy outcomes and sperm antibodies effect PSA.
+He has over eight years of experience in the Infertility field helping patients, jointly running a Andrology lab, and contributing research to advance the field through joint research on how AMH effects pregnancy outcomes and sperm antibodies effect PSA.
 He has a certificate in Data Analytics from Georgia Institute of Technology.

### 2022-02-19 paper-capstone-earthquake.tex
fx compile errors

@@
 \PassOptionsToPackage{hyphens}{url}
-\documentclass[sigplan,screen]{format/acmart}
+\documentclass[sigplan,screen]{acmart}
 \usepackage{todonotes}
@@
 \usepackage{ifthen}
-\usepackage[numbers]{natbib}‎
+% \usepackage[numbers]{natbib}‎
 
-\usepackage[colorlinks = true,
-            linkcolor = blue,
-            urlcolor  = blue,
-            citecolor = blue,
-            anchorcolor = blue]{hyperref}
+%\usepackage[colorlinks = true,
+%            linkcolor = blue,
+%            urlcolor  = blue,
+%            citecolor = blue,
+%            anchorcolor = blue]{hyperref}
 
@@
 
-%% Not sure what's the correct copyright to set in this template...  See the below link for
-%% the allowed options in Table 3.  Would this be acmlicensed as we're retaining copyright,
-%% but are allowing ACM's usage?  Or are we publishing elsewhere and this should be none?
-%% https://www.acm.org/binaries/content/assets/publications/consolidated-tex-template/acmart.pdf#page=18
 \setcopyright{none}
-%\setcopyright{Apache 2.0}
 %\setcopyright{rightsretained}
@@
 
-%\acmConference[T. Buttler, R. Knuuti, J, Kolesar, G. C. Fox,
-%  G. von Laszewski, J. Fox, University of Virginia Data Science
-%  Capstone 2022]{MLCommons Earthquake Science Benchmark}{Jan. 22 --
-%  Aug. 31 2022}{Charlotsville, VA, USA}
-
-\acmConference[Data Science
-  Capstone, University of Virginia, 2022]{MLCommons Earthquake Science Benchmark}{Jan. 22 --
-  Aug. 31 2022}{Charlotsville, VA, USA}
+\acmConference[Data Science Capstone, University of Virginia,
+  2022]{MLCommons Earthquake Science Benchmark}{Jan. 22 -- Aug. 31
+  2022}{Charlotsville, VA, USA}
 
@@
 
-%% see http://dl.acm.org/ccs.cfm.
 \begin{CCSXML}
@@
 \maketitle
+
+\TODO{Fix the classification. See  \url{see http://dl.acm.org/ccs.cfm}}
 
@@
          \ref{id:nersc} & NERSC  & Perlmutter & \cite{www-perlmutter} &         6000 & \nvidia{A100}      & & & & Jan 2022      \\ \hline
-         \ref{id:sdsc}  & SDSC   & Expanse & \cite{www-expanse}       &          200 & \nvidid{V100}
+         \ref{id:sdsc}  & SDSC   & Expanse & \cite{www-expanse}       &          200 & \nvidia{V100}
          SMX2 & & & &               \\ \hline
@@
             &    &  &       &            8 & \nvidia{A100}      & 40GB & 8 & &       \\
-            &    &  &       &            ? & \nvidid{RTX3090}   & 24GB & ? & ? & 2021          \\
+            &    &  &       &            ? & \nvidia{RTX3090}   & 24GB & ? & ? & 2021          \\
             &    &  &       &            ? & \nvidia{K80}       & 11GB & 8 & 9 & 2021          \\
@@
 
-{\bf Thomas Butler}
-is a Graduate Student at University of Virginia's School of Data Science.
-His undergraduate degree is in Biomedical engineering.
-He has over eight years of experience in the Infertility field helping patients, jointly running a Andrology lab, and contributing research to advance the field through joint research on how AMH effects pregnancy outcomes and sperm antibodies effect PSA.
-He has a certificate in Data Analytics from Georgia Institute of Technology.
+{\bf Thomas Butler} is a Graduate Student at University of Virginia's
+School of Data Science. His undergraduate degree is in Biomedical
+engineering. He has over eight years of experience in the Infertility
+field helping patients, jointly running a Andrology lab, and
+contributing research to advance the field through joint research on
+how AMH effects pregnancy outcomes and sperm antibodies effect PSA. He
+has a certificate in Data Analytics from Georgia Institute of
+Technology.
 
-{\bf Robert Knuuti}
-is a Graduate Student at University of Virginia's School of Data Science.
-He has over 10 years experience in system architecture and software engineering, and specializes in Development Operations and Cloud Computing.
-He has constructed air gapped Continuous Integration and Continuous Delivery systems for multiple organizations each supporting more than 100 developers and has facilitated the construction of repeatable, tractable builds for users of these systems.
+{\bf Robert Knuuti} is a Graduate Student at University of Virginia's
+School of Data Science. He has over 10 years experience in system
+architecture and software engineering, and specializes in Development
+Operations and Cloud Computing. He has constructed air gapped
+Continuous Integration and Continuous Delivery systems for multiple
+organizations each supporting more than 100 developers and has
+facilitated the construction of repeatable, tractable builds for users
+of these systems.
 
-{\bf Jake Kolessar}
-{is a Graduate Student at the University of Virginia's School of Data Science.
-He has a background in mechanical engineering and 2 years of experience as a Modeling, Simulation & Analysis Engineer.
-He has supported the software design and development of modeling capabilities for event simulation products as well as the integration of models into the simulation framework.}
+{\bf Jake Kolessar} {is a Graduate Student at the University of
+  Virginia's School of Data Science. He has a background in mechanical
+  engineering and 2 years of experience as a Modeling, Simulation and
+  Analysis Engineer. He has supported the software design and
+  development of modeling capabilities for event simulation products
+  as well as the integration of models into the simulation framework.}
 
@@
 
-{\bf Gregor von Laszewski}
-is a Research Professor at University of Virginia.
-He has more than 30 years of experience in parallel and distributed computing.
-Selected research organizations he worked in include NASA, Argonne National Laboratory, Indiana University.
-He is proud to have been contributing members of research teams that invented hybrid parallel genetic algorithms, Grid computing, and the first large scale academic hybrid cloud, as well as the larges scientific bibliometric analysis of XSEDE in the world.
+{\bf Gregor von Laszewski} is a Research Professor at University of
+Virginia. He has more than 30 years of experience in parallel and
+distributed computing. Selected research organizations he worked in
+include NASA, Argonne National Laboratory, Indiana University. He is
+proud to have been contributing members of research teams that
+invented hybrid parallel genetic algorithms, Grid computing, and the
+first large scale academic hybrid cloud, as well as the larges
+scientific bibliometric analysis of XSEDE in the world.
 
@@
 \subsection{Generated List}
-\listoftodos{}
 
@@
 
+\clearpage
+
+\listoftodos{}
+
 \end{document}

### 2022-02-19 paper-capstone-earthquake.tex
bio figures

@@
 \usepackage{ifthen}
-% \usepackage[numbers]{natbib}‎
-
-%\usepackage[colorlinks = true,
-%            linkcolor = blue,
-%            urlcolor  = blue,
-%            citecolor = blue,
-%            anchorcolor = blue]{hyperref}
 
@@
 \xpretocmd{\digraph}{\addFileDependency{#2.dot}}{}{}
+
+\newcommand{\GITISSUE}[1]{\href{https://github.com/Data-ScienceHub/mlcommons-science/issues/#1}{GitHub Issue #1}}}
+\newcommand{\GITPULL}[1]{\href{https://github.com/laszewsk/mlcommons/pull/#1}{GitHub Pull #1}}}
 
@@
 
-Several research efforts have been executed over the years to establish a complex set of factors that accurately predict where an earthquake could occur, and what impacts an earthquake would cause to the environment when they occur.
-Not only is this an algorithmically complex problem, but it is also computationally dense and requires high performance hardware to obtain these predictions\cite{fox2022aiforscience}.
-The intersection of these two elements create an ideal environment for building fair benchmarks.
-By establishing a rigorous process for constructing and running deep learning models the basis for comparing not just computational power in the High Performance Computing ecosystem, but also establish a benchmarks of new scientific discoveries from its execution.
+Several research efforts have been executed over the years to
+establish a complex set of factors that accurately predict where an
+earthquake could occur, and what impacts an earthquake would cause to
+the environment when they occur.  Not only is this an algorithmically
+complex problem, but it is also computationally dense and requires
+high performance hardware to obtain these
+predictions\cite{fox2022aiforscience}.  The intersection of these two
+elements create an ideal environment for building fair benchmarks.  By
+establishing a rigorous process for constructing and running deep
+learning models the basis for comparing not just computational power
+in the High Performance Computing ecosystem, but also establish a
+benchmarks of new scientific discoveries from its execution.
 
-For this study, we review the following model architectures across various hardware platforms to establish this benchmark:
+For this study, we review the following model architectures across
+various hardware platforms to establish this benchmark:
+
 \begin{itemize}
@@
 
-Each framework have their own performance measures that will shape the reference benchmark implementation.
+Each framework have their own performance measures that will shape the
+reference benchmark implementation.
 
@@
 
-MLCommons is an open engineering organization that seeks to unify the engineering and machine learning communities through system benchmarks, open data, and best practices \cite{www-mlcommons}.
-By using open data, a reference implementation for benchmarking, and executing best practices, the necessary rigor enables the construction repeatable research to establish a basis when performing future analysis.
-Using MLCommons organizational standards set by \mlperf{}, the dimensions of a the minimally viable solution for benchmarking are know and to establish our comparison criteria across our models and the executions on the different hardware sets\cite{mattson2019mlperf}.
+MLCommons is an open engineering organization that seeks to unify the
+engineering and machine learning communities through system
+benchmarks, open data, and best practices \cite{www-mlcommons}.  By
+using open data, a reference implementation for benchmarking, and
+executing best practices, the necessary rigor enables the construction
+repeatable research to establish a basis when performing future
+analysis.  Using MLCommons organizational standards set by \mlperf{},
+the dimensions of a the minimally viable solution for benchmarking are
+know and to establish our comparison criteria across our models and
+the executions on the different hardware sets\cite{mattson2019mlperf}.
 
@@
 
-The \mlperf{} framework already have several feature complete benchmarks that can judge system performance, however these benchmarks are centered on well known algorithms with solutions that are well known.
-Additionally, some CPU architectures create optimized instructions to solve common algorithmic problems at the microcode level, creating a performance boost when these prepared instructions are encountered\cite{cheong1997optimizer}.
-While these measures are still of great importance, they do not reflect the performance of systems based on non-deterministic measures that exist in nature.
-This makes benchmarking using a scientific dataset an ideal candidate for calculating performance, as there's an inherit randomness that's difficult to "game" this system and build predictions.
-This is especially true when attempting to predict the next occurrence of an earthquake, as complex models are required to gain predictive insights\cite{www-murakami2021}.
+The \mlperf{} framework already have several feature complete
+benchmarks that can judge system performance, however these benchmarks
+are centered on well known algorithms with solutions that are well
+known.  Additionally, some CPU architectures create optimized
+instructions to solve common algorithmic problems at the microcode
+level, creating a performance boost when these prepared instructions
+are encountered\cite{cheong1997optimizer}.  While these measures are
+still of great importance, they do not reflect the performance of
+systems based on non-deterministic measures that exist in nature.
+This makes benchmarking using a scientific dataset an ideal candidate
+for calculating performance, as there's an inherit randomness that's
+difficult to "game" this system and build predictions.  This is
+especially true when attempting to predict the next occurrence of an
+earthquake, as complex models are required to gain predictive
+insights\cite{www-murakami2021}.
 
@@
 
-MLCommons Science Research group is involved in edge and data center issues, end-to-end systems, inference and training.
-MLCommons Science Research focuses on Science applications and not industry.
-There are similarities between industry and science, for example image data is similar but some algorithms can be very different.
-The best example of this is particle physics experiments, quarks and lumens are not a common thing to model in industry \cite{www-mlcommons-science}.
+MLCommons Science Research group is involved in edge and data center
+issues, end-to-end systems, inference and training. MLCommons Science
+Research focuses on Science applications and not industry. There are
+similarities between industry and science, for example image data is
+similar but some algorithms can be very different. The best example of
+this is particle physics experiments, quarks and lumens are not a
+common thing to model in industry \cite{www-mlcommons-science}.
 
-The primary goal of MLCommons Science Research is to develop optimal methods to improve both speed and efficiency of scientific outcomes.
-This is done by measure how different models perform on different scientific datasets, in particular score the models based on how well they do on scientific discovery.
-This could be anywhere from accuracy of a model to time to reach  specific accuracy target when new data is introduced.
-In general, how well each model matches the goals of that particular scientific field.
-By using verified datasets following best standardized practice on how to build and exchange models using the reference materials on benchmarked machines, knowledge can flow more easily and freely between various researchers.
-By having these standard methods, both the monetary cost and time can be reduced between researcher regardless of their speciality.
+The primary goal of MLCommons Science Research is to develop optimal
+methods to improve both speed and efficiency of scientific
+outcomes. This is done by measure how different models perform on
+different scientific datasets, in particular score the models based on
+how well they do on scientific discovery. This could be anywhere from
+accuracy of a model to time to reach specific accuracy target when new
+data is introduced. In general, how well each model matches the goals
+of that particular scientific field. By using verified datasets
+following best standardized practice on how to build and exchange
+models using the reference materials on benchmarked machines,
+knowledge can flow more easily and freely between various
+researchers. By having these standard methods, both the monetary cost
+and time can be reduced between researcher regardless of their
+speciality.
 
-Currently MLCommons Science Research has four datasets based on climate, material science, medicine and earthquakes with a fifth dataset on plasma physics being added soon.
-Each dataset has their own method used to solve their scientific problem \cite{www-mlcommons-science}.
+Currently MLCommons Science Research has four datasets based on
+climate, material science, medicine and earthquakes with a fifth
+dataset on plasma physics being added soon. Each dataset has their own
+method used to solve their scientific problem
+\cite{www-mlcommons-science}.
 
@@
 
-\mlcube{} is  a set of standards and software developed by MLCommons for delivering machine learning models across container runtimes such as docker, kubernetes, singularity, or for automating the model's execution over ssh.
-\mlcube{} provides a common configuration format and interface such that consumers of a model do not need to worry about the underlying configuration and baseline of supporting libraries, and enables the user to focus on the model itself\cite{www-mlcube}.
+\mlcube{} is a set of standards and software developed by MLCommons
+for delivering machine learning models across container runtimes such
+as docker, kubernetes, singularity, or for automating the model's
+execution over ssh.  \mlcube{} provides a common configuration format
+and interface such that consumers of a model do not need to worry
+about the underlying configuration and baseline of supporting
+libraries, and enables the user to focus on the model
+itself\cite{www-mlcube}.
 
@@
 \begin{itemize}
-    \item \textbf{Developer.}  The developer produces a YAML configuration file that describes the target execution details including platform, runtime dependencies, inputs, outputs, metadata about the model, and the source code.
-Additionally, this configuration file offers branching paths based on what \mlcube{} runtime backend is selected (for example, the container image to use for docker can be configured differently than a kubeflow configuration).
-This is so the implementation details can be explicitly addressed, if necessary, by the model developer and abstracts this complexity away from consumers.
-    \item \textbf{User.}  The user of the system will receive the \mlcube{} code repository and execute a command that identifies which platform they intend to run the model on and any type of input parameters to the model detailed by the developer.
+
+    \item \textbf{Developer.}  The developer produces a YAML
+      configuration file that describes the target execution details
+      including platform, runtime dependencies, inputs, outputs,
+      metadata about the model, and the source code.  Additionally,
+      this configuration file offers branching paths based on what
+      \mlcube{} runtime backend is selected (for example, the
+      container image to use for docker can be configured differently
+      than a kubeflow configuration).  This is so the implementation
+      details can be explicitly addressed, if necessary, by the model
+      developer and abstracts this complexity away from consumers.
+
+    \item \textbf{User.}  The user of the system will receive the
+      \mlcube{} code repository and execute a command that identifies
+      which platform they intend to run the model on and any type of
+      input parameters to the model detailed by the developer.
 \end{itemize}
 
-As described by the \mlcube{} development team, this ecosystem establishes a contract between model developers and consumers of that model so that results can be easily shared and provide a framework to produce consistent results.
+As described by the \mlcube{} development team, this ecosystem
+establishes a contract between model developers and consumers of that
+model so that results can be easily shared and provide a framework to
+produce consistent results.
 
-The MLCommons Earthquake Science Benchmark models target the \mlcube{} mechanism for delivery.
+The MLCommons Earthquake Science Benchmark models target the \mlcube{}
+mechanism for delivery.
 
@@
 
-Earthquake prediction is a complex and old problem.
-There are two major tasks related to earthquakes.
-One is forecasting earthquake occurrences/likelihoods.
-Another is to predict damage once an earthquake does happen.
-The first is difficult because the details of the underground plates and the friction laws between them are not known and furthermore the earthquake causes phase transitions between plates which makes for unpredictable movement.
-The second, we need to predict earthquake energy wave movements.
-Since both of these problems are so complex with many hidden variables and the equations governing the phenomenon are unknown or incomplete, deep learning can be use to learn the various hidden patterns in the data which the model can then use to forecast earthquakes into the future.
-\cite{fox2021earthquake}
+Earthquake prediction is a complex and old problem. There are two
+major tasks related to earthquakes. One is forecasting earthquake
+occurrences/likelihoods. Another is to predict damage once an
+earthquake does happen. The first is difficult because the details of
+the underground plates and the friction laws between them are not
+known and furthermore the earthquake causes phase transitions between
+plates which makes for unpredictable movement. The second, we need to
+predict earthquake energy wave movements. Since both of these problems
+are so complex with many hidden variables and the equations governing
+the phenomenon are unknown or incomplete, deep learning can be use to
+learn the various hidden patterns in the data which the model can then
+use to forecast earthquakes into the future. \cite{fox2021earthquake}
 
@@
 
-\TODO{This section is in progress! Enter Energy Weighted Quantities equations picture below}
+\TODO{This section is in progress! Enter Energy Weighted Quantities
+  equations picture below}
 
-This function maximizes the largest earthquake energy values for the 2 week period. 
-Magnitude averaged over the bin
-All input values and predictions are independently normalized and have a maximum modulus of 1 over all space and time values. \cite{fox2022aiforscience}
+This function maximizes the largest earthquake energy values for the 2
+week period.  Magnitude averaged over the bin All input values and
+predictions are independently normalized and have a maximum modulus of
+1 over all space and time values. \cite{fox2022aiforscience}
 
@@
 
-The data being used is the Southern California location data, between 32 to 36 lat -120 to -114 long.
-The data is recorded between 1950-2020.
-Each location point is is 0.1 by 0.1 degrees or ll km by ll km in size or 2400 points in total.
-The points are then put into spatial bag which are based on a bunch of different location points, necessarily nearby spatially.
-Each location point has a time series associated with it.
-The time series data holds magnitude, depth, spatial position, and time.
-\cite{fox2021earthquake}
+The data being used is the Southern California location data, between
+32 to 36 lat -120 to -114 long. The data is recorded between
+1950-2020. Each location point is is 0.1 by 0.1 degrees or ll km by ll
+km in size or 2400 points in total. The points are then put into
+spatial bag which are based on a bunch of different location points,
+necessarily nearby spatially. Each location point has a time series
+associated with it. The time series data holds magnitude, depth,
+spatial position, and time. \cite{fox2021earthquake}
 
-The data was accumulated into 2 week samples using energy-based averaging, the log loss equation.
+The data was accumulated into 2 week samples using energy-based
+averaging, the log loss equation.
 
@@
 
-The Aurora supercomputer, sponsored by the U.S.
-Department of Energy (DOE), is part of the Argonne Leadership Computing Facility (ACLF).
-It was developed with the help of industry experts from Intel and Cray.
-Aurora offers HPC for data analytics and AI development with a peak performance of over 2 Exaflop DP.
-The system is based on the HPE Cray EX supercomputer platform and offers over 9,000 compute nodes.
+The Aurora supercomputer, sponsored by the U.S. Department of Energy
+(DOE), is part of the Argonne Leadership Computing Facility (ACLF). It
+was developed with the help of industry experts from Intel and
+Cray. Aurora offers HPC for data analytics and AI development with a
+peak performance of over 2 Exaflop DP. The system is based on the HPE
+Cray EX supercomputer platform and offers over 9,000 compute nodes.
 
@@
 
-The Summit supercomputer, sponsored by the U.S.
-Department of Energy (DOE), is located in the Oak Ridge Leadership Computing Facility (OLCF).
-The open science computing system is available to government, academic, and industry researchers to solve complex problems in all fields of science.
-The system is powered by 4,608 nodes which each house multiple IBM POWER9 CPUs and NVIDIA Volta GPUs on top of half a terabyte of coherent memory.
+The Summit supercomputer, sponsored by the U.S. Department of Energy
+(DOE), is located in the Oak Ridge Leadership Computing Facility
+(OLCF). The open science computing system is available to government,
+academic, and industry researchers to solve complex problems in all
+fields of science. The system is powered by 4,608 nodes which each
+house multiple IBM POWER9 CPUs and NVIDIA Volta GPUs on top of half a
+terabyte of coherent memory.
 
@@
 
-The Perlmutter supercomputer, sponsored by the U.S.
-Department of Energy (DOE), is located in Shyh Wang Hall at Berkeley Lab and is operated by the National Energy Research Scientific Computing Center (NERSC).
-The supercomputer has a 12 cabinet GPU systems with over 1,500 nodes and a 12 cabinet CPU system containing over 3,000 nodes.
-It is supported by 35 petabytes of all-flash storage which can move data at 5 terabytes/sec.
+The Perlmutter supercomputer, sponsored by the U.S. Department of
+Energy (DOE), is located in Shyh Wang Hall at Berkeley Lab and is
+operated by the National Energy Research Scientific Computing Center
+(NERSC). The supercomputer has a 12 cabinet GPU systems with over
+1,500 nodes and a 12 cabinet CPU system containing over 3,000
+nodes. It is supported by 35 petabytes of all-flash storage which can
+move data at 5 terabytes/sec.
 
@@
 UVA, Rivanna, A100, RTX3090, others may not have enough memory.
-\cite{www-rivanna}
-\TODO{description of Rivanna on web page is not accurate.}
+\cite{www-rivanna} \TODO{description of Rivanna on web page is not
+  accurate.}
 
-Containers are available on Rivanna through Singularity \cite{www-singularity} that can be loaded with modules \cite{www-modules}.
-We plan to utilize singularity, but target the formulation of the container in docker and than transform it to singularity.
-As singularity is installed also on other machines, we will leverage this deployment also on them.
+Containers are available on Rivanna through Singularity
+\cite{www-singularity} that can be loaded with modules
+\cite{www-modules}. We plan to utilize singularity, but target the
+formulation of the container in docker and than transform it to
+singularity. As singularity is installed also on other machines, we
+will leverage this deployment also on them.
 
@@
 
-The PEARL system, funded by the Alan Turing Institute, is hosted at the Rutherford Appleton Laboratory by the Science and Technology Facilities Council (STFC).
-It was developed for the advancement of AI and Machine Learning research with large-scale scientific datasets, already contributing to the fields of materials, environmental and life sciences, and astronomy.
-It is comprised of two NVIDIA DGX2 nodes, 32 GPUs, 3TB of system memory, 1TB of GPU memory, and 600TB of high-speed storage.
+The PEARL system, funded by the Alan Turing Institute, is hosted at
+the Rutherford Appleton Laboratory by the Science and Technology
+Facilities Council (STFC). It was developed for the advancement of AI
+and Machine Learning research with large-scale scientific datasets,
+already contributing to the fields of materials, environmental and
+life sciences, and astronomy. It is comprised of two NVIDIA DGX2
+nodes, 32 GPUs, 3TB of system memory, 1TB of GPU memory, and 600TB of
+high-speed storage.
 
@@
 
-\TODO{wrapfigure does somehow not work in this template}
+\newcommand{\BIOFIG}[3]{%
 
-{\bf Thomas Butler} is a Graduate Student at University of Virginia's
-School of Data Science. His undergraduate degree is in Biomedical
-engineering. He has over eight years of experience in the Infertility
-field helping patients, jointly running a Andrology lab, and
-contributing research to advance the field through joint research on
-how AMH effects pregnancy outcomes and sperm antibodies effect PSA. He
-has a certificate in Data Analytics from Georgia Institute of
-Technology.
+  {\bf #2} #3
+  %\begin{wrapfigure}[l]{.2\columnwidth}
+  \includegraphics[width=.2\columnwidth]{#1}%
+  %\end{wrapfigure}
+}
 
-{\bf Robert Knuuti} is a Graduate Student at University of Virginia's
-School of Data Science. He has over 10 years experience in system
-architecture and software engineering, and specializes in Development
-Operations and Cloud Computing. He has constructed air gapped
-Continuous Integration and Continuous Delivery systems for multiple
-organizations each supporting more than 100 developers and has
-facilitated the construction of repeatable, tractable builds for users
-of these systems.
+\BIOFIG{images/bio/gregor.png}{Thomas Butler}{is a Graduate Student at
+  University of Virginia's School of Data Science. His undergraduate
+  degree is in Biomedical engineering. He has over eight years of
+  experience in the Infertility field helping patients, jointly
+  running a Andrology lab, and contributing research to advance the
+  field through joint research on how AMH effects pregnancy outcomes
+  and sperm antibodies effect PSA. He has a certificate in Data
+  Analytics from Georgia Institute of Technology.}
 
-{\bf Jake Kolessar} {is a Graduate Student at the University of
-  Virginia's School of Data Science. He has a background in mechanical
-  engineering and 2 years of experience as a Modeling, Simulation and
-  Analysis Engineer. He has supported the software design and
-  development of modeling capabilities for event simulation products
-  as well as the integration of models into the simulation framework.}
+\BIOFIG{images/bio/gregor.png}{Robert Knuuti}{is a Graduate Student at
+  University of Virginia's School of Data Science. He has over 10
+  years experience in system architecture and software engineering,
+  and specializes in Development Operations and Cloud Computing. He
+  has constructed air gapped Continuous Integration and Continuous
+  Delivery systems for multiple organizations each supporting more
+  than 100 developers and has facilitated the construction of
+  repeatable, tractable builds for users of these systems.}
 
-{\bf Geoffrey C. Fox}
-\TODO{Geoffery: fill out}
+\BIOFIG{images/bio/gregor.png}{Jake Kolessar}{is a Graduate Student at
+  the University of Virginia's School of Data Science. He has a
+  background in mechanical engineering and 2 years of experience as a
+  Modeling, Simulation and Analysis Engineer. He has supported the
+  software design and development of modeling capabilities for event
+  simulation products as well as the integration of models into the
+  simulation framework.}
+
+\BIOFIG{images/bio/gregor.png}{Geoffrey C. Fox}{TBD TBD TBD TBD TBD
+  TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD
+  TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD
+  TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD
+  TBD TBD TBD TBD TBD TBD TBD TBD}
 
@@
 
-{\bf Gregor von Laszewski} is a Research Professor at University of
-Virginia. He has more than 30 years of experience in parallel and
-distributed computing. Selected research organizations he worked in
-include NASA, Argonne National Laboratory, Indiana University. He is
-proud to have been contributing members of research teams that
-invented hybrid parallel genetic algorithms, Grid computing, and the
-first large scale academic hybrid cloud, as well as the larges
-scientific bibliometric analysis of XSEDE in the world.
+\BIOFIG{images/bio/gregor.png}{Gregor von Laszewski}{is a Research
+  Professor at University of Virginia.  He has more than 30 years of
+  experience in parallel and distributed computing.  Selected research
+  organizations he worked at include NASA, Argonne National
+  Laboratory, Indiana University. He is proud to have been
+  contributing members of research teams that invented hybrid parallel
+  genetic algorithms, Grid computing, and the first large scale
+  academic hybrid cloud, as well as the larges scientific bibliometric
+  analysis of XSEDE in the world.}
 
@@
     \item Learn how to selfdeploy jupyterlab \cite{www-jupyterlab}
-    \item Begin to learn about papermill \cite{www-papermill} \TODO{Rivanna's core configurations might be a bit limited on this front.
-They do provide multiple versions of software (such as py2.7 and py3.8), but we may need to get permission to do something in userspace if we need to be very specific on a version of python.
-Do we want to look into using tox, conda, or leave this up to the container ecosystem to solve?}
+
+    \item Begin to learn about papermill \cite{www-papermill}
+      \TODO{Rivanna's core configurations might be a bit limited on
+        this front.  They do provide multiple versions of software
+        (such as py2.7 and py3.8), but we may need to get permission
+        to do something in userspace if we need to be very specific on
+        a version of python.  Do we want to look into using tox,
+        conda, or leave this up to the container ecosystem to solve?}
+
       \TODO{Gregor: add how to use conda and modules to switch python versions}
@@
     \item depends on Tensorflow
-    \TODO{Below this line are objectives or targets to take the current modeling solution and mature it for other platforms / ecosystems.
-Rivanna uses Lua's lmod ecosystem for jailing a process, and anaconda uses solved environments for dependencies that extend beyond just python modules.}
+
+    \TODO{Below this line are objectives or targets to take the
+      current modeling solution and mature it for other platforms /
+      ecosystems.  Rivanna uses Lua's lmod ecosystem for jailing a
+      process, and anaconda uses solved environments for dependencies
+      that extend beyond just python modules.}
+
   \item Familiarize yourself with Rivanna's modules \cite{www-modules}  and conda \cite{www-conda} environments.
@@
     \item horovod \cite{www-horovod}
-    \TODO{MLCommons project that can target a few platforms using a YAML contract.  Once we solve the target environment, we can likely target porting this way.}
+
+      \TODO{MLCommons project that can target a few platforms using a
+        YAML contract.  Once we solve the target environment, we can
+        likely target porting this way.}
 
@@
 \begin{itemize}
-\item Reviewed background information \footnote{\url{https://github.com/Data-ScienceHub/mlcommons-science/issues/1}}
+\item Reviewed background information \GITISSUE{1}
     \begin{itemize}
@@
     \end{itemize}
-\item Reviewed background technologies \footnote{\url{https://github.com/Data-ScienceHub/mlcommons-science/issues/2}}
+\item Reviewed background technologies \GITISSUE{2}
     \begin{itemize}
@@
     \end{itemize}
-\item Revised Intro to Python guide for conda installations so it doesn't interfere with pyenv or other shell commands.\footnote{\url{https://github.com/cloudmesh-community/book/pull/574}} \footnote{\url{https://github.com/cloudmesh-community/book/pull/575}}
+
+  \item Revised Intro to Python guide for conda installations so it
+    doesn't interfere with pyenv or other shell
+    commands.\footnote{\url{https://github.com/cloudmesh-community/book/pull/574}} \footnote{\url{https://github.com/cloudmesh-community/book/pull/575}}
+
   \item Solved conda environment and requirements.txt for colab notebook
-\footnote{\url{https://github.com/Data-ScienceHub/mlcommons-science/issues/12}} \footnote{\url{https://github.com/laszewsk/mlcommons/pull/1}} \footnote{\url{https://github.com/laszewsk/mlcommons/pull/3}}
+
+    \GITISSUE{12},
+    \GITPULL{1},
+    \GITPULL{3}.
+
 \item build docker container image that can run python 3.9.7 and conda
 \item verified access to GPU allocations in Rivanna
+
 \end{itemize}

### 2022-02-19 paper-capstone-earthquake.tex
bio images
finish bio image
todo added
fix latex syntax error

@@
 \providecommand{\mlcube}{MLCube}
+
+\newcommand{\BIOFIG}[3]{%
+\parindent 0pt%
+\setlength{\intextsep}{6pt}%
+\setlength{\columnsep}{6pt}%
+
+{\bf #2}
+
+\begin{wrapfigure}{r}{0.2\columnwidth}%
+    \includegraphics[width=0.2\columnwidth]{#1}%
+\end{wrapfigure}%
+#3%
+}
 
@@
 
-\newcommand{\GITISSUE}[1]{\href{https://github.com/Data-ScienceHub/mlcommons-science/issues/#1}{GitHub Issue #1}}}
-\newcommand{\GITPULL}[1]{\href{https://github.com/laszewsk/mlcommons/pull/#1}{GitHub Pull #1}}}
+\newcommand{\GITISSUE}[1]{\href{https://github.com/Data-ScienceHub/mlcommons-science/issues/#1}{GitHub Issue #1}}
+\newcommand{\GITPULL}[1]{\href{https://github.com/laszewsk/mlcommons/pull/#1}{GitHub Pull #1}}
 
@@
 
-The NVIDIA DGX Station A100 is a commercially available product that can provide powerful computing resources to all industries.
-The DGX Station is an AI resource that can run parallel jobs and support training, inference, and data analytics.
-It contains 4 NVIDIA A100 GPUs and up to 320 GBs of GPU memory.
-With the use of Multi-Instance GPU (MIG), each A100 GPU can be partitioned into 7 isolated instances.
+The NVIDIA DGX Station A100 is a commercially available product that
+can provide powerful computing resources to all industries. The DGX
+Station is an AI resource that can run parallel jobs and support
+training, inference, and data analytics. It contains 4 NVIDIA A100
+GPUs and up to 320 GBs of GPU memory. With the use of Multi-Instance
+GPU (MIG), each A100 GPU can be partitioned into 7 isolated instances.
 
@@
 
-\newcommand{\BIOFIG}[3]{%
-
-  {\bf #2} #3
-  %\begin{wrapfigure}[l]{.2\columnwidth}
-  \includegraphics[width=.2\columnwidth]{#1}%
-  %\end{wrapfigure}
-}
-
 \BIOFIG{images/bio/gregor.png}{Thomas Butler}{is a Graduate Student at
@@
 
-\BIOFIG{images/bio/gregor.png}{Geoffrey C. Fox}{TBD TBD TBD TBD TBD
+\BIOFIG{images/bio/fox.png}{Geoffrey C. Fox}{TBD TBD TBD TBD TBD
   TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD
@@
 \end{itemize}
+
+\TODO{There are more we developed in the web page}
 

### 2022-02-19 paper-capstone-earthquake.tex
rewrite of hardware

@@
 
-\subsection{Aurora}
+\subsubsection{Aurora}
 \label{id:anl}
 
-The Aurora supercomputer, sponsored by the U.S. Department of Energy
-(DOE), is part of the Argonne Leadership Computing Facility (ACLF). It
-was developed with the help of industry experts from Intel and
-Cray. Aurora offers HPC for data analytics and AI development with a
-peak performance of over 2 Exaflop DP. The system is based on the HPE
-Cray EX supercomputer platform and offers over 9,000 compute nodes.
+The Aurora supercomputer \cite{www-aurora} is locatd at that Argonne
+Leadership Computing Facility (ACLF) part of the U.S. Department of
+Energy (DOE).  It was developed with the help of industry experts from
+Intel and Cray. Aurora offers HPC for data analytics and AI
+development with a peak performance of over 2 Exaflop DP. The system
+is based on the HPE Cray EX supercomputer platform and offers over
+9,000 compute nodes.
 
-Argonne National Laboratory (ANL), ... \cite{www-aurora}
-
-\subsection{Summit}
+\subsubsection{Summit}
 \label{id:ornl}
 
-The Summit supercomputer, sponsored by the U.S. Department of Energy
-(DOE), is located in the Oak Ridge Leadership Computing Facility
-(OLCF). The open science computing system is available to government,
-academic, and industry researchers to solve complex problems in all
-fields of science. The system is powered by 4,608 nodes which each
-house multiple IBM POWER9 CPUs and NVIDIA Volta GPUs on top of half a
+The Summit supercomputer\cite{www-summit} is located in the Oak Ridge
+Leadership Computing Facility (OLCF) part of the U.S. Department of
+Energy (DOE). The system is avalable to government, academic, and
+industry researchers to solve complex problems in all fields of
+science. The system is powered by 4,608 nodes which each house
+multiple IBM POWER9 CPUs and NVIDIA Volta GPUs on top of half a
 terabyte of coherent memory.
 
-Oak Ridge National Laboratory (ORNL), Summit \cite{www-summit}
-
-\subsection{Perlmutter}
+\subsubsection{Perlmutter}
 \label{id:nersc}
 
-The Perlmutter supercomputer, sponsored by the U.S. Department of
-Energy (DOE), is located in Shyh Wang Hall at Berkeley Lab and is
-operated by the National Energy Research Scientific Computing Center
-(NERSC). The supercomputer has a 12 cabinet GPU systems with over
-1,500 nodes and a 12 cabinet CPU system containing over 3,000
-nodes. It is supported by 35 petabytes of all-flash storage which can
-move data at 5 terabytes/sec.
+The Perlmutter \cite{www-perlmutter} supercomputer is located at the
+National Energy Research Scientific Computing Center (NERSC) which is
+part of the U.S. Department of Energy (DOE).  It has a 12 cabinet GPU
+systems with over 1,500 nodes and a 12 cabinet CPU system containing
+over 3,000 nodes \TODO{This is inconsistant or not properly explained
+  as 2 node numbers are used}. It is supported by 35 petabytes of
+all-flash storage which can move data at 5 terabytes/sec.
 
-National Energy Research Scientific Computing Center (NERSC), ...
-\cite{www-perlmutter}
+\subsubsection{Expanse}
+\label{id:sdsc}
 
-\subsection{SDSC}
-\label{id:sdsc} 
-\TODO{Which of the SDSC systems will we be benchmarking?}
-San Diego Super Computing Center, ... \cite{www-expanse}
+The Expanse \cite{www-expanse} system at the Sandiego Super Computing
+Center (SDSC) ....
 
-\subsection{University of Virginia Rivanna}
+\TODO{Expanse description incomplete}
+
+\subsubsection{Rivanna}
 \label{id:uva}
+
+University of Virginia Rivanna
+
+\TODO{Rivann description is inomplete}
 
@@
 
-The PEARL system, funded by the Alan Turing Institute, is hosted at
-the Rutherford Appleton Laboratory by the Science and Technology
-Facilities Council (STFC). It was developed for the advancement of AI
-and Machine Learning research with large-scale scientific datasets,
-already contributing to the fields of materials, environmental and
-life sciences, and astronomy. It is comprised of two NVIDIA DGX2
-nodes, 32 GPUs, 3TB of system memory, 1TB of GPU memory, and 600TB of
-high-speed storage.
-
-Rutherford Appleton Laboratory (RAL), UK 
-Pearl \cite{www-pearl-1}
+The PEARL \cite{www-pearl-1} system is located t the Alan Turing
+Institute in the the Rutherford Appleton Laboratory, UK.  It was
+developed for the advancement of AI and Machine Learning research with
+large-scale scientific datasets, already contributing to the fields of
+materials, environmental and life sciences, and astronomy. It is
+comprised of two NVIDIA DGX2 nodes, 32 GPUs, 3TB of system memory, 1TB
+of GPU memory, and 600TB of high-speed storage.
 
@@
 
-The NVIDIA DGX Station A100 is a commercially available product that
-can provide powerful computing resources to all industries. The DGX
-Station is an AI resource that can run parallel jobs and support
-training, inference, and data analytics. It contains 4 NVIDIA A100
-GPUs and up to 320 GBs of GPU memory. With the use of Multi-Instance
-GPU (MIG), each A100 GPU can be partitioned into 7 isolated instances.
-
-DGX Fox \cite{www-dgx-station-a100}
+The NVIDIA DGX Station A100 \cite{www-dgx-station-a100} is a
+commercially available product that can provide powerful computing
+resources to all industries. The DGX Station is an AI resource that
+can run parallel jobs and support training, inference, and data
+analytics. It contains 4 NVIDIA A100 GPUs and up to 320 GBs of GPU
+memory. With the use of Multi-Instance GPU (MIG), each A100 GPU can be
+partitioned into 7 isolated instances.
 

### 2022-02-19 paper-capstone-earthquake.tex
update hw table for desktop

@@
     \caption{Overview of compute resources.}
-    \label{tab:my_label}
+    \label{tab:hwoverview}
     \centering
@@
         \hline
-        Section & Organization   & Machine                         & & Processors    & GPUs             & \makecell{Memory\\/Device} &  \makecell{GPU\\/Device} & \makecell{No.
+        See also & Organization   & Machine                         & & Processors    & GPUs             & \makecell{Memory\\/Device} &  \makecell{GPU\\/Device} & \makecell{No.
 of\\ nodes} & Commissioned \\
@@
     \caption{Overview of compute resources.}
-    \label{tab:my_label}
+    \label{tab:desktops}
     \centering
@@
         \hline
-        Section & Organization   & Machine                         & Processors    & GPUs             & \makecell{Memory\\/Device} &  \makecell{GPU\\/Device} & \makecell{No.
+        See also & Organization   & Machine                         & Processors    & GPUs             & \makecell{Memory\\/Device} &  \makecell{GPU\\/Device} & \makecell{No.
 of\\ nodes} & Commissioned \\
@@
         \hline
-         \ref{id:gcomputer}   & 
+         Table \ref{tab:gcompute}   & 
          Gregor    & 
@@
          \hline
-         \ref{id:tcomputer}   & Thomas.A  & i7-8750H &  1 &   NVIDIA GeForce RTX 2070 with Max-Q Design  & 16GB & 1 & 1 & 2019   \\
+         \ref{sec:tcomputer}   & Thomas.A  & i7-8750H &  1 &   NVIDIA GeForce RTX 2070 with Max-Q Design  & 16GB & 1 & 1 & 2019   \\
          \hline
-         \ref{id:t2computer}   & Thomas.B  & i7-7700HQ &  1 &   NVIDIA GeForce GTX 1060  & 16GB & 1 & 1 & 2017   \\
+         \ref{sec:t2computer}   & Thomas.B  & i7-7700HQ &  1 &   NVIDIA GeForce GTX 1060  & 16GB & 1 & 1 & 2017   \\
          \hline
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\input{machines}
 

### 2022-02-19 bin/sysinfo.py

@@ -1,0 +1,6 @@
+from cloudmesh.common.StopWatch import StopWatch
+
+StopWatch.start("dummy")
+StopWatch.stop("dummy")
+
+StopWatch.benchmark()

### 2022-02-21 paper-capstone-earthquake.tex
Updates from Overleaf
citations

@@
 \usepackage{ifthen}
+\usepackage{comment}
 
@@
 
-\acmBooktitle{Woodstock '18:
-  ACM Symposium on Neural Gaze Detection, June 03--05, 2018,
-  Woodstock, NY}
+\acmBooktitle{Report for the MLCommons Science Working Group in collaboration with University of Virginia, May 2022, Charlottesville, VA}
 
@@
 \TODO{Fix the classification. See  \url{see http://dl.acm.org/ccs.cfm}}
+
+\begin{comment}
+\parindent 0pt {\bf Citation:}
+
+\begin{footnotesize}
+\begin{verbatim}@TechReport{mlcommons-eartquake,
+  author = 	 {Thomas Butler and Robert Knuuti and
+              Jake Kolessar and Geoffrey C. Fox and
+              Gregor von Laszewski and Judy Fox},
+  title = 	 {MLCommons Earthquake Science Benchmark},
+  institution =  {MLCommons Science Working Group},
+  year = 	 2022,
+  type = 	 {Report by University of Virginia},
+  address = 	 {Charlottesville, VA},
+  month = 	 may,
+  note = 	 {The order of the authors and url location may change},
+  annote = 	 {Version: draft},
+  url = {https://github.com/cyberaide/paper-capstone-mlcommons}
+}\end{verbatim}
+\end{footnotesize}
+\end{comment}
 
@@
          \hline
+         Table \ref{tab:rcompute} & Rob.K & AMD Ryzen 9 5900HX & 1 &  NVidia RTX 3080 &  16GB & 1 & 1 & 11/25/2021 \\
+         \hline
     \end{tabular}
@@
 \end{table*}
+
+\input{machines}
 
@@
 
-\BIOFIG{images/bio/gregor.png}{Robert Knuuti}{is a Graduate Student at
+\BIOFIG{images/bio/rob.png}{Robert Knuuti}{is a Graduate Student at
   University of Virginia's School of Data Science. He has over 10

### 2022-02-21 paper-capstone-earthquake.tex
add markdown

@@
 \usepackage{ifthen}
+% \usepackage[smartEllipses]{markdown}
 \usepackage{comment}
@@
 
+\newcommand*\rot{\rotatebox{90}}
 \newcommand{\BIOFIG}[3]{%
@@
 \begin{CCSXML}
-<ccs2012>
+<mdcs2012>
  <concept>
@@
 \begin{adjustbox}{max width=\textwidth}
-    \begin{tabular}{|r|l|l|r|l|l|l|l|l|}
+    \begin{tabular}{|r|l|l|l|r|r|r|l|r|l|l|l|}
         \hline
-        See also & Organization   & Machine                         & Processors    & GPUs             & \makecell{Memory\\/Device} &  \makecell{GPU\\/Device} & \makecell{No.
-of\\ nodes} & Commissioned \\
+        See also & 
+        Organization   & 
+        Machine                         &
+        OS & 
+        \rot{Processors~} &
+        \rot{Core~} & 
+        \rot{Threads~} & 
+        GPUs             & 
+        \rot{\makecell{Memory\\/Device}} &  
+        \rot{GPU/Device} & 
+        \rot{No. of nodes~} & 
+        Commissioned \\ 
         \hline
@@
          5950X   &  
+         Ubuntu 20.04&
          1 &   
+         16 & 
+         32 &
          \href{https://www.gigabyte.com/Graphics-Card/GV-N307TGAMING-OC-8GD#kf}{GIGABYTE Gaming RTX 3070TI}  & 
          8GB & 
-         1 & 
+         2 &
          1 &
@@
             &
+            & 
+            &  
+            &   
          \href{https://rog.asus.com/us/graphics-cards/graphics-cards/rog-strix/rog-strix-rtx3090-o24g-gaming-model/}{ASUS ROG Strix RTX 3090} & 
          24GB & 
+         & 
+         & 
+         Feb 2022   \\
+         \hline
+         \ref{sec:tcomputer}   & Thomas.A  & i7-8750H & Win11 Pro  & 1 &   ? & ? & NVIDIA GeForce RTX 2070 with Max-Q Design  & 16GB & 1 & 1 & 2019   \\
+         \hline
+         \ref{sec:t2computer}   & Thomas.B  & i7-7700HQ &  Win11 Pro &1 & ? & ? &  NVIDIA GeForce GTX 1060  & 16GB & 1 & 1 & 2017   \\
+         \hline
+         Table \ref{tab:rcompute} & 
+         Rob & 
+         5900HX & 
+         Win11 Pro & 
+         1 & 
+         8 & 
+         16 & 
+         NVIDIA RTX 3080 &  
+         16GB & 
          1 & 
          1 & 
-         Feb 2022   \\
-         \hline
-         \ref{sec:tcomputer}   & Thomas.A  & i7-8750H &  1 &   NVIDIA GeForce RTX 2070 with Max-Q Design  & 16GB & 1 & 1 & 2019   \\
-         \hline
-         \ref{sec:t2computer}   & Thomas.B  & i7-7700HQ &  1 &   NVIDIA GeForce GTX 1060  & 16GB & 1 & 1 & 2017   \\
-         \hline
-         Table \ref{tab:rcompute} & Rob.K & AMD Ryzen 9 5900HX & 1 &  NVidia RTX 3080 &  16GB & 1 & 1 & 11/25/2021 \\
+         11/25/2021 \\
          \hline
@@
 
-\begin{acks}
-\TODO{Geoffrey: please add acknowledgement for funding, other acks}
-\end{acks}
-
 %
@@
 \input{machines}
+
+\begin{acks}
+\TODO{Geoffrey: please add acknowledgement for funding, other acks}
+\end{acks}
 
@@
 \begin{itemize}
-    \item Introduction to Python \cite{las-intro-python}
+    \item Introduction to Python \cite{las-intro-python}, Gregor von Laszewski, with contributions from Robert Knuuti
+    \item \href{https://laszewsk.github.io/mlcommons/docs/tutorials/rivanna/}{Running GPU Batch jobs on Rivanna} Gregor von Laszewski, Robert Knuuti
 \end{itemize}
@@
 \TODO{There are more we developed in the web page}
+
+% \markdownInput{rivanna-manual}
 

### 2022-02-24 paper-capstone-earthquake.tex
Updates from Overleaf

--- 2022-02-21 paper-capstone-earthquake.tex
@@
 
+\TODO{provide introduction to earthquake}
+
+\TODO{Review how these papers contribute to introduction}
+
+\begin{itemize}
+    \item Review IEEE Presentation \cite{fox2022aiforscience}
+    \item Review Fox paper \cite{fox2021earthquake}
+\end{itemize}
+
+\section{Earthquake Data}
+
 \subsection{Energy-based averaging, the log loss equation}
@@
 averaging, the log loss equation.
-
-\TODO{provide introduction to earthquake}
-
-\TODO{Review how these papers contribute to introduction}
-
-\begin{itemize}
-    \item Review IEEE Presentation \cite{fox2022aiforscience}
-    \item Review Fox paper \cite{fox2021earthquake}
-\end{itemize}
 
@@
          \hline
-         \ref{sec:tcomputer}   & Thomas.A  & i7-8750H &  1 &   NVIDIA GeForce RTX 2070 with Max-Q Design  & 16GB & 1 & 1 & 2019   \\
+         Table \ref{tab:tcompute}   & Thomas.A  & i7-8750H &  1 & \href{https://www.gigabyte.com/Laptop/AERO-15--RTX-20-Series#kf}{NVIDIA GeForce RTX 2070 with Max-Q Design} & 16GB & 1 & 1 & Jan 2019   \\
          \hline
-         \ref{sec:t2computer}   & Thomas.B  & i7-7700HQ &  1 &   NVIDIA GeForce GTX 1060  & 16GB & 1 & 1 & 2017   \\
-         \hline
-         Table \ref{tab:rcompute} & Rob.K & AMD Ryzen 9 5900HX & 1 &  NVidia RTX 3080 &  16GB & 1 & 1 & 11/25/2021 \\
+         Table \ref{tab:rcompute} & Rob.K & AMD Ryzen 9 5900HX & 1 &  \href{https://rog.asus.com/us/laptops/rog-strix/2021-rog-strix-scar-15-series/}{NVidia RTX 3080} &  16GB & 1 & 1 & 11/25/2021 \\
          \hline
@@
 
-\input{machines}
-
 \bibliographystyle{ACM-Reference-Format}
@@
 
-\BIOFIG{images/bio/gregor.png}{Thomas Butler}{is a Graduate Student at
+\BIOFIG{images/bio/tom.png}{Thomas Butler}{is a Graduate Student at
   University of Virginia's School of Data Science. His undergraduate
@@
 
-\BIOFIG{images/bio/fox.png}{Geoffrey C. Fox}{TBD TBD TBD TBD TBD
-  TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD
-  TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD
-  TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD
-  TBD TBD TBD TBD TBD TBD TBD TBD}
+\BIOFIG{images/bio/fox.png}{Geoffrey C. Fox}{is a computational and data 
+  science professor for University of Virginia. He has over 50 years of 
+  experience in 'big data' and high-performance computing. Through his 
+  career, he has contributed to these fields in both research and education 
+  the next generation. In 2019 he received the ACM-IEEE CS Ken Kennedy 
+  award for his work in parallel computing methodology, algorithms and 
+  software, data analysis, and how these interface with broad classes of 
+  applications. His current research interests are exploring the various 
+  use cases of neural networks.}
 
@@
 \item verified access to GPU allocations in Rivanna
+\item created a portable version of python 3.10.2 that can run on Rivanna \GITPULL{8}
+\item implemented cloudmesh-data (\href{https://github.com/cloudmesh/cloudmesh-data/pull/1}{cloudmesh/cloudmesh-data PR 1}, \href{https://github.com/cloudmesh/cloudmesh-data/pull/3}{cloudmesh/cloudmesh-data PR #3})
 

### 2022-02-25 paper-capstone-earthquake.tex

--- 2022-02-21 paper-capstone-earthquake.tex
@@
          \hline
-         \ref{sec:tcomputer}   & Thomas.A  & i7-8750H & Win11 Pro  & 1 &   ? & ? & NVIDIA GeForce RTX 2070 with Max-Q Design  & 16GB & 1 & 1 & 2019   \\
+         \ref{sec:tcomputer}   & Thomas.A  & i7-8750H & Win11 Pro  & 1 &   ?
+& ?
+& \href{https://www.gigabyte.com/Laptop/AERO-15--RTX-20-Series#kf}{NVIDIA GeForce RTX 2070 with Max-Q Design}NVIDIA GeForce RTX 2070 with Max-Q Design  & 16GB & 1 & 1 & 2019   \\
          \hline
@@
          16 & 
-         NVIDIA RTX 3080 &  
+         \href{https://rog.asus.com/us/laptops/rog-strix/2021-rog-strix-scar-15-series/}{NVidia RTX 3080} &  
          16GB & 
@@
 
-\input{machines}
-
 \begin{acks}

### 2022-02-25 paper-capstone-earthquake.tex
Updates from Overleaf

--- 2022-02-24 paper-capstone-earthquake.tex
@@
 
-\subsection{Energy-based averaging, the log loss equation}
-
-\TODO{This section is in progress! Enter Energy Weighted Quantities
-  equations picture below}
-
-This function maximizes the largest earthquake energy values for the 2
-week period.  Magnitude averaged over the bin All input values and
-predictions are independently normalized and have a maximum modulus of
-1 over all space and time values. \cite{fox2022aiforscience}
-
 \subsection{Earthquake Data turned into Spatial bag Time Series}
@@
 associated with it. The time series data holds magnitude, depth,
-spatial position, and time. \cite{fox2021earthquake}
+spatial position, and time. The data was accumulated into 2 week samples. 
+\cite{fox2021earthquake}
 
-The data was accumulated into 2 week samples using energy-based
-averaging, the log loss equation.
+\subsection{Energy-based averaging, the log loss equation}
+
+\TODO{This section is in progress!}
+
+The data was filtered so only significant quakes were considered 
+(magnitude > 3.29). The Log Energy equation (Figure 1) was used for 
+the magnitude of significant quakes inside a 2 week period.
+
+\begin{figure}[htp]
+    \centering
+    \includegraphics[width=0.75\columnwidth]{images/Equations/log_energy.png}
+    \caption{Log Energy \cite{fox2021earthquake}}
+\end{figure}
+
+The Energy Weighting equation (Figure 2) was used to average depth from quakes.
+
+\begin{figure}[htp]
+    \centering
+    \includegraphics[width=0.75\columnwidth]{images/Equations/Q_bin.png}
+    \caption{Energy Weighted \cite{fox2021earthquake}}
+\end{figure}
+
+The Multiplicity equation (Figure 3) is used to count the number of 
+occurrences of significant earthquakes in a 2 week period.
+
+\begin{figure}[htp]
+    \centering
+    \includegraphics[width=0.75\columnwidth]{images/Equations/Multiplicity_bin.png}
+    \caption{Multiplicity \cite{fox2021earthquake}}
+\end{figure}
+
+This function maximizes the largest earthquake energy values for the 2
+week period.  Magnitude averaged over the bin, all input values and
+predictions are independently normalized and have a maximum modulus of
+1 over all space and time values. \cite{fox2022aiforscience}
 
@@
         Rivanna Frontend & 1m11.535s & 1m00.780s & 0m10.352s & N/A & N/A & N/A & & Intel(R) Xeon(R) CPU E5-2630 v3 @ 2.40GHz & Q3  2014 \\ \hline
+        Rob K & 
     \end{tabular}
@@
 
-\TODO{Rivann description is inomplete}
+\TODO{Rivanna description is incomplete}
 

### 2022-02-25 paper-capstone-earthquake.tex
update table

@@
          1 &
-         Sep 2021   \\
-            & 
-            &
-            &
-            &
-            & 
-            &  
-            &   
+         Sep 2021-Feb 2022   \\
+         \hline
+         Table \ref{tab:gcompute}   & 
+         Gregor    & 
+         5950X   &  
+         Ubuntu 20.04&
+         1 &   
+         16 & 
+         32 &
          \href{https://rog.asus.com/us/graphics-cards/graphics-cards/rog-strix/rog-strix-rtx3090-o24g-gaming-model/}{ASUS ROG Strix RTX 3090} & 
@@
 & ?
-& \href{https://www.gigabyte.com/Laptop/AERO-15--RTX-20-Series#kf}{NVIDIA GeForce RTX 2070 with Max-Q Design}NVIDIA GeForce RTX 2070 with Max-Q Design  & 16GB & 1 & 1 & 2019   \\
+& \href{https://www.gigabyte.com/Laptop/AERO-15--RTX-20-Series#kf}{NVIDIA GeForce RTX 2070 with Max-Q Design}  & 16GB & 1 & 1 & 2019   \\
          \hline

### 2022-03-04 paper-capstone-earthquake.tex
fault groupings, accuracy metrics, spacial ..., averaging data equations

@@
 systems based on non-deterministic measures that exist in nature.
-This makes benchmarking using a scientific dataset an ideal candidate
+This makes benchmarking using a scientific dataframe an ideal candidate
 for calculating performance, as there's an inherit randomness that's
@@
 outcomes. This is done by measure how different models perform on
-different scientific datasets, in particular score the models based on
+different scientific dataframes, in particular score the models based on
 how well they do on scientific discovery. This could be anywhere from
@@
 data is introduced. In general, how well each model matches the goals
-of that particular scientific field. By using verified datasets
+of that particular scientific field. By using verified dataframes
 following best standardized practice on how to build and exchange
@@
 
-Currently MLCommons Science Research has four datasets based on
+Currently MLCommons Science Research has four dataframes based on
 climate, material science, medicine and earthquakes with a fifth
-dataset on plasma physics being added soon. Each dataset has their own
+dataframe on plasma physics being added soon. Each dataframe has their own
 method used to solve their scientific problem
@@
 
-\subsection{Earthquake Data turned into Spatial bag Time Series}
-
-\TODO{This section is in progress!}
+\subsection{Spatial bag Time Series}
 
@@
 
-\subsection{Energy-based averaging, the log loss equation}
+\subsection{Averaging Data equations}
 
-\TODO{This section is in progress!}
-
-The data was filtered so only significant quakes were considered 
-(magnitude > 3.29). The Log Energy equation (Figure 1) was used for 
-the magnitude of significant quakes inside a 2 week period.
+The data was filtered so only quakes with a magnitude > 3.29, or significant 
+quakes, were considered. The Log Energy equation (Figure 1) was used for 
+the magnitude of significant quakes inside a 2 week period. The Log Energy 
+equation maximizes the largest earthquake energy values for the 2 week period. 
+\cite{fox2021earthquake}
 
@@
 
-This function maximizes the largest earthquake energy values for the 2
-week period.  Magnitude averaged over the bin, all input values and
-predictions are independently normalized and have a maximum modulus of
-1 over all space and time values. \cite{fox2022aiforscience}
+All input values and predictions are independently normalized and have a maximum 
+modulus of 1 over all space and time values. \cite{fox2022aiforscience}
+
+\subsection{Fault Groupings}
+
+\TODO{Work in Progress}
+
+Grouped faults empirically into 36 families by proximity. #Show picture of groupings. Used 4 different space filling curves to fill in empty spaces to get graph into a 2d structure.
 
@@
 % \end{figure}
+
+\subsection{Accuracy Metrics}
+
+\TODO{either put plots here or reference where they are}
+
+The models are measured according to two accuracy metrics, the Nash Sutcliffe Efficiency 
+NSE and a visualization of predicted verse observed earthquake data. NSE measures 
+the variation difference between observed and predicted data over time. $Q^t_m$ is the 
+model prediction at time t, $Q^t_o$ is observed value at time t, $\overline{Q_o}$ is the 
+mean observed value over all time. \cite{fox2021earthquake}
+
+\begin{figure}[htp]
+    \centering
+    \includegraphics[width=0.75\columnwidth]{images/Equations/NSE_equation.png}
+    \caption{Nash Sutcliffe Efficiency \cite{fox2021earthquake}}
+\end{figure}
+
+The results are presented as the normalized NSE or NNSE. NNSE = 1/(2-NSE)
+The values of NNSE ranges between 0 and 1 with 0 = mean, the higher the value the better the fit. \cite{fox2021earthquake}
 

### 2022-10-16 paper-capstone-earthquake.tex
Updates from Overleaf

@@
 \documentclass[sigplan,screen]{acmart}
+
+\usepackage{amsmath}
 \usepackage{todonotes}
@@
 The data being used is the Southern California location data, between
-32 to 36 lat -120 to -114 long. The data is recorded between
-1950-2020. Each location point is is 0.1 by 0.1 degrees or ll km by ll
-km in size or 2400 points in total. The points are then put into
+$32$ to $36$ lat $-120$ to $-114$ long. The data is recorded between
+$1950-2020$. Each location point is $0.1$ by $0.1$ degrees or $11 km$ by $11
+km$ in size or $2400$ points in total. The points are then put into
 spatial bag which are based on a bunch of different location points,
@@
 associated with it. The time series data holds magnitude, depth,
-spatial position, and time. The data was accumulated into 2 week samples. 
+spatial position, and time. The data was accumulated into $2$ week samples. 
 \cite{fox2021earthquake}
@@
 
-The data was filtered so only quakes with a magnitude > 3.29, or significant 
-quakes, were considered. The Log Energy equation (Figure 1) was used for 
-the magnitude of significant quakes inside a 2 week period. The Log Energy 
-equation maximizes the largest earthquake energy values for the 2 week period. 
-\cite{fox2021earthquake}
+The data was filtered so only quakes with a $magnitude > 3.29$, or significant 
+quakes, were considered. The Log Energy equation (\ref{eq:Log Energy}) was used for 
+the magnitude of significant quakes inside a $2$ week period. The Log Energy 
+equation maximizes the largest earthquake energy values for the $2$ week period
+\cite{fox2021earthquake}.
 
-\begin{figure}[htp]
-    \centering
-    \includegraphics[width=0.75\columnwidth]{images/Equations/log_energy.png}
-    \caption{Log Energy \cite{fox2021earthquake}}
-\end{figure}
+    \begin{equation}
+    m_{bin} = \log (Energy) = \frac{1}{1.5} \log_{10} \sum\limits^{Events}_{in\ bin} 10^{1.5m_{event}}
+    \label{eq:Log Energy}
+    \end{equation}
 
-The Energy Weighting equation (Figure 2) was used to average depth from quakes.
+\TODO{Gregor suggests to change the event to e(1.5m) Gregor's suggestion just to show it see \ref{eq:log-energy-gregor}.
+possibly we should use  $in\ bin(1.5m)$ and than just $10^e$ ?.
+However we need to confirm.}
 
-\begin{figure}[htp]
-    \centering
-    \includegraphics[width=0.75\columnwidth]{images/Equations/Q_bin.png}
-    \caption{Energy Weighted \cite{fox2021earthquake}}
-\end{figure}
+    \begin{equation}
+    m_{bin} = \log (Energy) = \frac{1}{1.5} \log_{10} \sum\limits^{Events}_{e \in bin} 10^{e(1.5m)}
+    \label{eq:log-energy-gregor}
+    \end{equation}
 
-The Multiplicity equation (Figure 3) is used to count the number of 
-occurrences of significant earthquakes in a 2 week period.
+    \begin{equation}
+    m_{bin} = \log (Energy) = \frac{1}{1.5} \log_{10} \sum\limits^{Events}_{e \in bin(1.5m)} 10^e
+    \label{eq:log-energy-gregor}
+    \end{equation}
 
-\begin{figure}[htp]
-    \centering
-    \includegraphics[width=0.75\columnwidth]{images/Equations/Multiplicity_bin.png}
-    \caption{Multiplicity \cite{fox2021earthquake}}
-\end{figure}
+The Energy Weighting equation (\ref{eq:Energy Weighted}) was used to average depth from quakes \cite{fox2021earthquake}.
+
+    \begin{equation}
+    Energy\ weighted\ Quantity \ Q_{bin} = \dfrac{\sum\limits^{Events}_{in\ bin} 10^{1.5m_{event}}Q_{event}}{\sum\limits^{Events}_{in\ bin}10^{1.5m_{event}}}
+    \label{eq:Energy Weighted}
+    \end{equation}
+
+The Multiplicity equation (\ref{eq:Multiplicity}) is used to count the number of 
+occurrences of significant earthquakes in a 2 week period \cite{fox2021earthquake}.
+
+    \begin{equation}
+    Multiplicity_{bin} = \sum^{Events}_{in\ bin} Multiplicity_{event}
+    \label{eq:Multiplicity}
+    \end{equation}
 
 All input values and predictions are independently normalized and have a maximum 
-modulus of 1 over all space and time values. \cite{fox2022aiforscience}
+modulus of $1$ over all space and time values. \cite{fox2022aiforscience}
 
@@
 
-\TODO{Work in Progress}
+There are hundreds of faults in the $32$ to $36$ lat $-120$ to $-114$ long area.
+To simplify these faults, this area was taken as a $2d$ structure spit into $2400$ individual pixels.
+These pixels were grouped empirically into $36$ families by proximity using $4$ different space filling curves to fill in empty spaces.
 
-Grouped faults empirically into 36 families by proximity. #Show picture of groupings. Used 4 different space filling curves to fill in empty spaces to get graph into a 2d structure.
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{images/Fault_groupings.png}
+    \caption{Fault Regions}
+\end{figure}
+
+\subsection{Training - Test split}
+
+From the $2400$ pixels in the $2d$ image, the $500$ most active pixels were picked.
+These $500$ were picked according to the magnitude > $3.29$ cut.
+$400$ were used for training and $100$ for validation and testing.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{images/Training_test.png}
+    \caption{Training - Test Split: $400$ Training points are shown in red and $100$ test/validation points are shown in green}
+\end{figure}
 
@@
 The models are measured according to two accuracy metrics, the Nash Sutcliffe Efficiency 
-NSE and a visualization of predicted verse observed earthquake data. NSE measures 
+($NSE$) and a visualization of predicted verse observed earthquake data. $NSE$ measures 
 the variation difference between observed and predicted data over time. $Q^t_m$ is the 
-model prediction at time t, $Q^t_o$ is observed value at time t, $\overline{Q_o}$ is the 
-mean observed value over all time. \cite{fox2021earthquake}
+model prediction at time t, $Q^t_o$ is observed value at time $t$, $\overline{Q_o}$ is the 
+mean observed value over all time \cite{fox2021earthquake}.
 
-\begin{figure}[htp]
-    \centering
-    \includegraphics[width=0.75\columnwidth]{images/Equations/NSE_equation.png}
-    \caption{Nash Sutcliffe Efficiency \cite{fox2021earthquake}}
-\end{figure}
+    \begin{equation}
+    NSE = 1 - \dfrac{\sum\limits^{t=T}_{t=1}(Q^t_m-Q^t_o)^2}{\sum\limits^{t=T}_{t=1}(Q^t_o-\overline{Q_o})^2}
+    \label{eq:Nash Sutcliffe Efficiency}
+    \end{equation}
 
-The results are presented as the normalized NSE or NNSE. NNSE = 1/(2-NSE)
-The values of NNSE ranges between 0 and 1 with 0 = mean, the higher the value the better the fit. \cite{fox2021earthquake}
+The results are presented as the Normalized Nash Sutcliffe Efficiency ($NNSE$) \cite{fox2021earthquake}
+$NNSE. NNSE = 1/(2-NSE)$
+The values of $NNSE$ ranges between $0$ and $1$ with $0 = mean$, the higher the value the better the fit \cite{fox2021earthquake}.
 
@@
         Rivanna Frontend & 1m11.535s & 1m00.780s & 0m10.352s & N/A & N/A & N/A & & Intel(R) Xeon(R) CPU E5-2630 v3 @ 2.40GHz & Q3  2014 \\ \hline
-        Rob K & 
+        Rob K & \\ \hline
+        Thomas AERO 15 & 0m3.387s & 0m00.000s & 0m0.015s & 511.79 & 11.6 & NVIDIA RTX 2070 & June 1, 2019 & Intel Core i7-8750H @ 2.20GHz & Q2  2018 \\ \hline
     \end{tabular}
@@
 
-\TODO{Rivanna description is incomplete}
+\TODO{Rivanna description needs review}
 
 UVA, Rivanna, A100, RTX3090, others may not have enough memory.
-\cite{www-rivanna} \TODO{description of Rivanna on web page is not
-  accurate.}
+\cite{www-rivanna} \TODO{Write paragraph describing Rivanna compute cluster}
 
@@
 
+\section{GPU Analysis}
+
+Sample Text about GPU Event Times
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{images/GPU/event_times.png}
+    \caption{GPU Event Times: Nvidia V100 2 Epoch Experiment on Rivanna localscratch}
+\end{figure}
+
+Sample Text about Epoch Validation Figure
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{images/GPU/validation.png}
+    \caption{Epoch Validation: Nvidia V100 2 Epoch Experiment on Rivanna localscratch}
+\end{figure}
+
+Sample Text about kWh/Epoch
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{images/GPU/kWh_per_epoch.png}
+    \caption{Average kWh per Epoch: }
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
          \ref{id:uva}   & UVA    & Rivanna &\cite{www-rivanna}       &            8 & \nvidia{A100}      & 80GB & ? & &  Feb 2022      \\
-            &    &  &       &            8 & \nvidia{A100}      & 40GB & 8 & &       \\
+            &    &  &       &            8 & \nvidia{A100}      & 40GB & 8 & 2 &       \\
             &    &  &       &            ? & \nvidia{RTX3090}   & 24GB & ? & ? & 2021          \\
-            &    &  &       &            ? & \nvidia{K80}       & 11GB & 8 & 9 & 2021          \\
+            &    &  &       &            ? & \nvidia{K80}       & 11GB & 8 & 9 (8?) & 2021          \\
             &    &  &       &            ? & \nvidia{V100}      & 16GB & 4 & 1 & 2021          \\
@@
          8GB & 
-         2 &
+         1 &
          1 &
@@
          24GB & 
-         & 
-         & 
+         1 & 
+         1 & 
          Feb 2022   \\
          \hline
-         \ref{sec:tcomputer}   & Thomas.A  & i7-8750H & Win11 Pro  & 1 &   ?
-& ?
-& \href{https://www.gigabyte.com/Laptop/AERO-15--RTX-20-Series#kf}{NVIDIA GeForce RTX 2070 with Max-Q Design}  & 16GB & 1 & 1 & 2019   \\
+         \ref{sec:tcomputer}   & Thomas.A  & i7-8750H & Win10 Pro  & 1 &   6 & 12 & \href{https://www.gigabyte.com/Laptop/AERO-15--RTX-20-Series#kf}{NVIDIA GeForce RTX 2070 with Max-Q Design}  & 16GB & 1 & 1 & 2019   \\
          \hline
-         \ref{sec:t2computer}   & Thomas.B  & i7-7700HQ &  Win11 Pro &1 & ? & ? &  NVIDIA GeForce GTX 1060  & 16GB & 1 & 1 & 2017   \\
+         \ref{sec:t2computer}   & Thomas.B  & i7-7700HQ &  Win10 Pro &1 & 4 & 8 &  NVIDIA GeForce GTX 1060  & 16GB & 1 & 1 & 2017   \\
          \hline
@@
 
-\BIOFIG{images/bio/gregor.png}{Jake Kolessar}{is a Graduate Student at
+\BIOFIG{images/bio/jake.jpg}{Jake Kolessar}{is a Graduate Student at
   the University of Virginia's School of Data Science. He has a
@@
   analysis of XSEDE in the world.}
+
+\BIOFIG{images/bio/j-fox.jpg}{Judy Fox} is an Associate Professor of Data  
+  Science and holds a joint appointment as Associate Professor with the
+  Department of Computer Science. Her research focuses on designing
+  experimental systems that enable Data Science, Machine Learning, and  the
+  Internet  of  Things  applications  to  harness  the  computational
+  resources  of  Cloud  and  HPC  platforms  effectively.
 
@@
     \item Become familiar with the Attention Paper \cite{vaswani2017attention}
-    \item Learn how to selfdeploy jupyterlab \cite{www-jupyterlab}
-
-    \item Begin to learn about papermill \cite{www-papermill}
-      \TODO{Rivanna's core configurations might be a bit limited on
-        this front.  They do provide multiple versions of software
-        (such as py2.7 and py3.8), but we may need to get permission
-        to do something in userspace if we need to be very specific on
-        a version of python.  Do we want to look into using tox,
-        conda, or leave this up to the container ecosystem to solve?}
 
